---
title: "AreWeRepresented Census and NCACC Categorical Data"
output: 
  html_document: default
author: "D. Hopp"
date: "April 15, 2018"
---

INPUT:  
AWR_Census_NCACC.RData  
df_nc_voted.RData  

OUTPUT:  
df_merge_county_sex_race_20161108 as a csv file  
AWR_Census_NCACC_cat.RData an amalgamation of AWR_Census_NCACC.RData and objects created in this script  

## Who is Using Which Codes for What  

The AWR_Census_NCACC.RData image includes the directory paths data_dir and output_dir  

```{r garb}
load("AWR_Census_NCACC.RData")
load("df_nc_voted.RData")
```  

## NCSBE  
### Registration Data Race and Ethnicity   

results_pct_20161108.txt via  
https://dl.ncsbe.gov/index.html?prefix=ENRS/2016_11_08/    

Defined as:  
dfcl_regis_party_cd <- df_regis_stats_20161108 %>% dplyr::distinct(party_cd)
dfcl_regis_race_code <- df_regis_stats_20161108 %>% dplyr::distinct(race_code)  
dfcl_regis_ethnic_code <- df_regis_stats_20161108 %>% dplyr::distinct(ethnic_code)  
dfcl_regis_sex_code <- df_regis_stats_20161108 %>% dplyr::distinct(sex_code)  

```{r rsvp}
dfcl_regis_party_cd
dfcl_regis_race_code
dfcl_regis_ethnic_code
dfcl_regis_sex_code
```  
Convenience data frame:  
Race (only) by county in wide format:  

```{r myth}
fn_df_regis_stats_20161108_county_race_wide <- function() {
  df_ <-
  df_regis_stats_20161108 %>%
  dplyr::select(
    total_voters,
    race_code,
    FIPS3
  ) %>%
  reshape2::dcast(
    FIPS3~race_code,
    value.var = "total_voters",
    sum
  )
  #
  return(df_)
}
```  

Convenience data frame:  
Race (only) by county in long format:  

```{r froth}
fn_df_regis_stats_20161108_county_race_long <- function(){
  df_ <-
  df_regis_stats_20161108 %>%
  dplyr::select(
    total_voters,
    race_code,
    FIPS3
  ) %>%
  dplyr::group_by(
    FIPS3,
    race_code
  ) %>%
  dplyr::summarize(
    N=sum(total_voters)
  ) %>%
  dplyr::ungroup()
  #
  return(df_)
}
```  

```{r calm}
str(fn_df_regis_stats_20161108_county_race_long())
```  

Race and sex by county:  

```{r ward}
fn_df_regis_stats_20161108_county_sex_race <- function(){
  df_ <-
  df_regis_stats_20161108 %>%
  dplyr::select(
    total_voters,
    sex_code,
    race_code,
    FIPS3
  ) %>% 
  dplyr::group_by(
    FIPS3,
    sex_code,
    race_code
  ) %>% 
  dplyr::summarise(
    N=sum(total_voters)
  ) %>%
  dplyr::ungroup()
  #
  return(df_)
}
```  

```{r rhino}
str(fn_df_regis_stats_20161108_county_sex_race())
```  

## Census Bureau  
### Race and Ethnicity  

cc-est2016-alldata-37.csv via  
https://www.census.gov/data/tables/2016/demo/popest/counties-detail.html  
"Annual County Resident Population Estimates by Age, Sex, Race, and Hispanic Origin: April 1, 2010 to July 1, 2016"  

Census 2010 Race Methodology: 
[here](https://www2.census.gov/programs-surveys/popest/technical-documentation/methodology/modified-race-summary-file-method/mrsf2010.pdf) 
and [here](https://www.census.gov/prod/cen2010/briefs/c2010br-02.pdf)  

Gender is for each Race, and also total, e.g.,  
TOT_MALE, TOT_FEMALE, WA_MALE, WA_FEMALE etc.  

WA White alone  
BA Black or African American alone  
IA American Indian and Alaska Native alone  
AA Asian alone  
NA Native Hawaiian and Other Pacific Islander alone  
TOM  Two or More Races  

Hispanic is double-counted but included as additional columns:  
H Hispanic  

These are double-counted and are not included:  
WAC  White alone or in combination  
BAC  Black or African American alone  
IAC  American Indian and Alaska Native alone or in combination  

Counts by total population (all) and by voting age (18 and older)  
Breakdown by race, ignoring ethnicity.  
This is for entire state:  

```{r tern}
fn_df_cc_est2016_tot_subset() %>%
dplyr::select(TOT_POP,nWA,nBA,nIA,nAA,nTOM,pop) %>%
reshape2::melt(
  id.vars="pop"
) %>%
reshape2::dcast(
  pop~variable,
  value.var = "value",
  sum
)
```  

This is by county (by FIPS3): df_cc_est2016_county_race_dcast  

```{r mulch}
str(df_cc_est2016_county_race_dcast)
```  

## Reasonable Mappings  

Gender:  
NCSBE sex_code: M, F, U
Census Gender is for each Race, and also total, e.g.,  TOT_MALE, TOT_FEMALE, WA_MALE, WA_FEMALE etc.  

Race:
NCSBE race_code: A, B, I, M, O, U, W (does not include Native Hawaiian explicitly)  
Census Race (see Gender): WA, BA, IA (American Indian), AA, NA (Native Hawaiian etc.), TOM (two or more)  

### Remapping  

Convince census to look like NCSBE data:  

```{r maria}
df_cc_est2016_county_sex_race_long_recoded <-
  df_cc_est2016_tot %>%
  dplyr::select( 
    FIPS3,
    WA_MALE,
    WA_FEMALE,
    BA_MALE,
    BA_FEMALE,
    IA_MALE,
    IA_FEMALE,
    AA_MALE,
    AA_FEMALE,
    NA_MALE,
    NA_FEMALE,
    TOM_MALE,
    TOM_FEMALE,
    pop
  ) %>%
  reshape2::melt(
    id.vars=c("pop","FIPS3")
  ) %>%
  dplyr::mutate(
    variable=as.character(variable)
  ) %>%
  dplyr::mutate(
    sex=
      ifelse(
        stringr::str_detect(variable,"FEMALE"),"F","M"
      ),
    race=
      ifelse(stringr::str_detect(variable,"TOT_"),"all",
        (ifelse(stringr::str_detect(variable,"WA_"),"W",        
          (ifelse(stringr::str_detect(variable,"BA_"),"B",
            (ifelse(stringr::str_detect(variable,"IA_"),"I",
              (ifelse(stringr::str_detect(variable,"AA_"),"A",
                (ifelse(stringr::str_detect(variable,"TOM_"),"M","O")
                      )
                    )
                  )
                )
              )
            )
          )
        ) 
      )
    )
  ) %>%
  dplyr::select(
    pop,
    FIPS3,
    sex,
    race,
    value
  ) %>%
  dplyr::rename(
    N=value
  ) %>%
  dplyr::arrange(
    pop,
    FIPS3,
    sex,
    race
  )
```   

```{r abba}
str(df_cc_est2016_county_sex_race_long_recoded)
```  

## NCACC    

Notice that categories do NOT provide sex vs. race, just totals.  
In that sense, NCACC data is not compatible with census or NCSBE.  
Keep in mind that race most likely is self-reported.  

```{r churn}
fn_df_NCACC_2016_demog_subset <- function(){
  df_ <-
  df_NCACC_2016_demog %>%
  dplyr::mutate(
    nmale=ncomm-nfemale,
    nother_race=ncomm-(nwhite+nafam+namin)
  ) %>%
  dplyr::select(
    FIPS3,
    nmale,
    nfemale,
    nwhite,
    nafam,
    namin,
    nother_race
  )
  #
  return(df_)
}
```  

```{r polka}
str(fn_df_NCACC_2016_demog_subset())
```  

## NCSBE Voting Data  

```{r burma}
df_nc_voted_subset_long <-
df_nc_voted %>%
dplyr::filter(
  status_cd=="A"
) %>%
dplyr::inner_join(
  df_FIPS %>%
    dplyr::select(
      county_id,
      FIPS3
    ),
  by="county_id"
) %>%
dplyr::select(
  FIPS3,
  gender_code,
  race_code
) %>%
reshape2::dcast(
  FIPS3+gender_code+race_code~.
) %>%
plyr::rename(
  c(
    "."="N",
    "race_code"="race",
    "gender_code"="sex"
  )
) %>%
na.omit() %>%
dplyr::arrange(
  FIPS3,
  sex,
  race
)
```  

```{r yodel}
str(df_nc_voted_subset_long)
```  

## Bring Together  

Create one data frame to hold all of above (long shape)  
* Columns:  
* pop: 
  + all: Census Bureau all population  
  + voteage: Census Bureau 18 years and older  
  + regis: NCSBE registered to vote  
  + voted: NCSBE voted  
* sex:  
  + F: Female  
  + M: Male  
  + U: Unknown  
* race:  
  A: Asian  
  B: African-American  
  I: American Indian  
  M: More than one  
  O: Other  
  U: Unknown  
  W: White  

```{r groton}
df_merge_county_sex_race_20161108 <-
  rbind(
    fn_df_regis_stats_20161108_county_sex_race() %>%
    dplyr::rename(
      sex=sex_code,
      race=race_code
    ) %>%
    dplyr::mutate(
      pop="regis"
    ) %>%
    dplyr::select(
      pop,
      FIPS3,
      sex,
      race,
      N
    ),
    df_cc_est2016_county_sex_race_long_recoded %>%
      dplyr::filter(
        race != "all"
      ),
    df_nc_voted_subset_long %>%
    dplyr::mutate(
      pop="voted"
    ) %>%
    dplyr::select(
      pop,
      FIPS3,
      sex,
      race,
      N
    )
  )
```  

```{r icy}
str(df_merge_county_sex_race_20161108)
```  

Write df_merge_county_sex_race_20161108 as a csv file.  

```{r pencil}
readr::write_csv(
  df_merge_county_sex_race_20161108,
  paste0(output_dir,"df_merge_county_sex_race_20161108.csv.gz")
)
```  

Any analysis has to be aware of the numbers by race,  
which present comparison challenges across the sources.  
The "O" and "U" categories are of particular note.  

O (Other) and U (Unknown) are used in the NCSBE data, as per
ncvhis_ncvoter_data_format.txt but only O in the Census data.  
Are both O designations to be considered comparable?  

Display counts by race:  

```{r spire}
fn_merge_county_sex_race_20161108_dcast <- function(){
  # all this to get a nice lookin dcast
  # there must be a better way
  # use an ordered factor
  #
  df_ <- df_merge_county_sex_race_20161108 
  df_$pop <- as.factor(df_$pop)
  df_$pop <- ordered(
    df_$pop, 
    levels=c("all","voteage","regis","voted")
  )
  #
  df_dcast <-
  df_ %>% 
  dplyr::select(
    pop,
    race,
    N
  ) %>% 
  reshape2::dcast(
    race~pop, 
    value.var="N",
    sum,  
    margins="race"
  )
  #
  return(
    df_dcast
  )
}
```    

```{r etch}
fn_merge_county_sex_race_20161108_dcast()
```  

## Finish  

Do not resave df_nc_voted  

```{r moose}
rm(df_nc_voted)
ls()
```

```{r flop}
save.image("AWR_Census_NCACC_cat.RData")
```  

### END  