---
title: "AreWeRepresented NCSBE Data V2"
output: 
  html_document: default
author: "D. Hopp"
date: "August 22, 2018"
--- 

Original version stable June 10, 2018.  

### Note on the Voter Registration Form  

The form is at https://www.ncsbe.gov/Portals/0/Forms/NCVoterRegForm06W.pdf 

Gender: Male, Female  
Ethnicity: Hispanic/Latino, Not ...  
Race: African American/Black, American Indian/Alaskan Native, Asian, Multiracial, Native Hawaiian/Pacific Islander, White, Other  

### R packages; Input/Output files etc.  

R requires packages plyr, dplyr, reshape2, lubridate, data.table, readr  

INPUT:  
FIPS.csv  
DATA FILES that must be present (see context for sources)  
Change as appropriate for your data:  
ncvhis_Statewide_20180331_11082016.txt  
ncvhis_Statewide_20180331_11082016.txt  (<- an extract from ncvhis_Statewide.txt)  

SIGNIFICANT R OBJECTS CREATED:  
df_FIPS  county and FIPS codes  
df_ncvoter_in  munged ncvoter  
df_ncvhis_in  munged ncvhis  
df_nc_voted  joined ncvoter and ncvhis  
df_ncvhis_dups  if any dups in ncvhis  reason_cd

OUTPUT:  
df_ncvoter_in.RData  
df_ncvhis_in.RData  
df_nc_voted.RData  
AWR_NCSBE.RData (image but not including df_ncvoter_in, df_ncvhis_in, df_nc_voted)  

DIRECTORY STRUCTURE:  

```{r yawn}
data_dir <- "data/"
data_ncvhis_dir <- "data/ncvhis/"
data_ncvoter_dir <- "data/ncvoter/"
output_dir <- "output/"
```  

Data files and as_of date:

```{r zebra}
#
# must use a canonical form for the date:
as_of <- "2018-04-07"
#
ncvhis_file <- "ncvhis_Statewide_20180407_11082016.txt"
ncvoter_file <- "ncvoter_Statewide_20180407.txt" 
ncvoter_stats_file <- "voter_stats_20161108.txt"
```  
Check for existence  

```{r same}
tmp_ <- paste0(data_ncvhis_dir,ncvhis_file)
if (!file.exists(tmp_)) {
  print(paste(tmp_,"does not exist."))
  knitr::knit_exit()
}
#
tmp_ <- paste0(data_ncvoter_dir,ncvoter_file)
if (!file.exists(tmp_)) {
  print(paste(tmp_,"does not exist."))
  knitr::knit_exit()
}
#
tmp_ <- paste0(data_dir,ncvoter_stats_file)
if (!file.exists(tmp_)) {
  print(paste(tmp_,"does not exist."))
  knitr::knit_exit()
}
rm(tmp_)
```  

## A. FIPS Code  

Add 3-char FIPS3 code since some sources use as a numeric, some as character.  
Also add county_id used in NCSBE data.  
This code chunk appears in several places to assure that df_FIPS exists.  

```{r wheel}
df_FIPS <- read.csv(
  paste0(data_dir,"FIPS.csv"),
  header=TRUE,
  stringsAsFactors = FALSE
) %>%
dplyr::rowwise() %>%
dplyr::mutate(
   FIPS3=paste0(substr("00",1,3-nchar(as.character(FIPS))),as.character(FIPS))
) %>%
as.data.frame() 
#
df_FIPS %<>%
cbind(
  rownames(df_FIPS)
) %>%
plyr::rename(
  c("rownames(df_FIPS)"="county_id")
) %>%
dplyr::mutate(
  county_id=as.character(county_id)
)
```  

```{r pray}
str(df_FIPS)
#
readr::write_csv(
  df_FIPS,
  paste0(output_dir,"df_FIPS.csv.gz")
)
```  

## B. REGISTERED VOTERS, NOT ACTUAL VOTES.  

This file rolls up voter characteristics into county/precinct/race/ethnicity/sex/age counts. There are four age groups:  
18-25, 26-40, 41-65, and over 66  

Further below are counts of various categories.  

Source: https://dl.ncsbe.gov/index.html?prefix=ENRS/2016_11_08/  

See https://vt.ncsbe.gov/RegStat/Results/?date=11%2F08%2F2016 for one-line county rollup.  

```{r slow}
df_regis_stats_20161108 <- read.csv(  
  paste0(data_dir,"voter_stats_20161108.txt"),
  sep="\t",
  header=TRUE,
  stringsAsFactors=FALSE
)
```  

```{r loud}
str(df_regis_stats_20161108)
```  

stat_type: `r unique(df_regis_stats_20161108$stats_type)`

voter_type is "voter" only  
update_date has no content  
replace any missing character-type with NA   

```{r hefty}
df_regis_stats_20161108 %<>%
  dplyr::select(
    -stats_type,
    -update_date
  )  %>%
  mutate_if(
    is.character, funs(trimws(.)) 
  ) %>%
  base::replace(., .=="", NA)
```  

```{r lisp}
str(df_regis_stats_20161108)
```  

Number of records with race_code NA: `r nrow(df_regis_stats_20161108 %>% dplyr::filter(is.na(race_code)))`  

Replace race_code NA with "U"

```{r fda}
df_regis_stats_20161108 %<>%
  dplyr::mutate(
    race_code=ifelse(is.na(race_code),"U",race_code)
  )
```  

```{r knee}
str(df_regis_stats_20161108)
```  

add FIPS3 and remove county_desc 

```{r isle}
df_regis_stats_20161108 %<>%
  dplyr::inner_join(
    df_FIPS %>% 
      dplyr::select(
        County,
        FIPS3
      ) %>%
      dplyr::mutate(
        County=toupper(County)
      ),
    by=c("county_desc"="County")
  ) %>%
  dplyr::select(
    -county_desc
  )
```  

```{r kazoo}
str(df_regis_stats_20161108)
```  

Note here that precinct_abbrv is missing in some records (WHY?!):  
The number of rows with missing data: `r sum(!complete.cases(df_regis_stats_20161108))`    
The number of rows with misssing precinct_abbrv: `r nrow(df_regis_stats_20161108[is.na(df_regis_stats_20161108$precinct_abbrv),])`  
  
Party codes: `r unique(df_regis_stats_20161108$party_cd)`  
Race codes: `r unique(df_regis_stats_20161108$race_code)`  
Ethnic codes: `r unique(df_regis_stats_20161108$ethnic_code)`  
Sex codes: `r unique(df_regis_stats_20161108$sex_code)`  

Create codelists:

```{r bah}
dfcl_regis_party_cd <- df_regis_stats_20161108 %>% dplyr::distinct(party_cd)
dfcl_regis_race_code <- df_regis_stats_20161108 %>% dplyr::distinct(race_code)
dfcl_regis_ethnic_code <- df_regis_stats_20161108 %>% dplyr::distinct(ethnic_code)
dfcl_regis_sex_code <- df_regis_stats_20161108 %>% dplyr::distinct(sex_code)
```  

```{r doubt}
str(dfcl_regis_party_cd)
str(dfcl_regis_race_code)
str(dfcl_regis_ethnic_code)
str(dfcl_regis_sex_code)
#
persons_df_regis_stats_20161108 <- sum(df_regis_stats_20161108$total_voters)
```  

Total number of registered persons: `r persons_df_regis_stats_20161108`  

Number of voters in age groups:

```{r brace}
df_regis_age_dcast <-
reshape2::dcast(
  df_regis_stats_20161108,
  age~.,
  value.var="total_voters",
  function(x) sum(x),
  margins=TRUE
)
```  

```{r sorry}
str(df_regis_age_dcast)
```  

### County rollup  
Ignore age.  
This dataframe facilitates summaries for party, sex, race, etc.  

```{r dixie}
df_regis_stats_20161108_county_rollup <-
  df_regis_stats_20161108 %>%
  dplyr::select(
    -precinct_abbrv,
    -election_date,
    -vtd_abbrv
  ) %>%
  dplyr::group_by(
    FIPS3,
    party_cd,
    race_code,
    ethnic_code,
    sex_code
  ) %>%
  dplyr::summarize(
    tot_voters=sum(total_voters)
  ) %>%
  dplyr::ungroup()
```  

Check on race categories:  

```{r warren}
(df_regis_stats_20161108_county_rollup %>% 
  reshape2::dcast(
    race_code~.,
    value.var="tot_voters",
    sum,
    margins=TRUE
  )
)
```  

A note on how to roll this up even tighter.  
This will create a data frame without party_cd or ethnic_code, 
and with totals adjusted accordingly.   

```{r alive}
fn_df_regis_stats_20161108_county_rollup_narrow <- function(){
  df_  <-
  df_regis_stats_20161108_county_rollup %>%
  dplyr::select(
    FIPS3,
    race_code,
    sex_code,
    tot_voters
  ) %>%
  dplyr::group_by(
    FIPS3,
    race_code,
    sex_code
  ) %>% 
  dplyr::summarise(
    N=sum(tot_voters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    FIPS3,
    N,
    sex_code,
    race_code
  )
  #
  return(df_)
}
```  

```{r banal}
str(fn_df_regis_stats_20161108_county_rollup_narrow())
```  

## C. ncvhis  

### Read ncvhis files and move selected data to postgreSQL data base.   

Ref: http://amunategui.github.io/dealing-with-large-files/   

### Postgres table def:  
  
Note: election is kept as a char field even though it seems to hold a data mm/dd/yyyy  
since there do not seem to be any guarantees that it will always be a date.  
Any sorts can be done in some pragmatic way.  
  
CREATE TABLE public.ncvhis  
(  
    county_id character(3) COLLATE pg_catalog."default" NOT NULL,  
    voter_reg_num character(12) COLLATE pg_catalog."default" NOT NULL,  
    election character(10) COLLATE pg_catalog."default" NOT NULL,  
    as_of_date date NOT NULL,  
    CONSTRAINT ncvhis_pkey PRIMARY KEY (county_id, voter_reg_num)  
)  
WITH (  
    OIDS = FALSE  
)  
TABLESPACE pg_default;  
  
ALTER TABLE public.ncvhis  
    OWNER to postgres;  
COMMENT ON TABLE public.ncvhis   
    IS 'from NCSBE ncvhis zip file';  

### Handling the data file ncvhis_Statewide.zip  

Use the ncvhis_Statewide.zip file located in data/ncvhis/chunk  
      IMPORTANT  
Prior to this script, unzip and use such as Windows FINDSTR to create a  
file with only the "11/08/2016" records:  
findstr /L "11/08/2016" ncvhis_Statewide.txt > ncvhis_Statewide_found.txt  
  
This has 4,768,160 lines   


```{r deter}
#
# FILE COLS:
#  "county_id","county_desc","voter_reg_num","election_lbl","election_desc",
#  "voting_method","voted_party_cd","voted_party_desc","pct_label",
#	"pct_description","ncid","voted_county_id","voted_county_desc","vtd_label",
#  "vtd_description"
#
# keep only these columns
#
file_cols_keep <- c(
  "county_id",
  "voter_reg_num",
  "election_lbl"
)
```  

### IMPORTANT:  
input txt file MUST ALREADY be filtered on election date  
fast read using data.table  
convert to data frame for ease of computation  

```{r one}
df_ncvhis_in <- data.table::fread(
  paste0(data_ncvhis_dir,ncvhis_file),
  header=FALSE,
  select=c(1,3,4),
  col.names=file_cols_keep
) %>%
as.data.frame() %>%
dplyr::rename(
  election=election_lbl
) %<>%
dplyr::mutate(
  as_of_date=lubridate::as_date(as_of)
)
```  

There may be duplicates  
if there are, remember them and remove them  

```{r land}
str(df_ncvhis_in)
```  

```{r lab}
# remove df_ncvhis_dups if it already exists
if (exists("df_ncvhis_dups")) {
  rm(df_ncvhis_dups)
}
#
v_ncvhis_dups <- 
  which(
    duplicated(
      df_ncvhis_in
    )
  )  
#
if (length(v_ncvhis_dups)>0) {
  # save list of dups 
  # (there are only 2 per instance although this might require verification in the future)
  df_ncvhis_dups <- 
    df_ncvhis_in %>% .[v_ncvhis_dups,]
  # remove them
  df_ncvhis_in %<>%
    dplyr::distinct(
    )
}
```  

```{r heat}
if (exists("df_ncvhis_dups")) {
  str(df_ncvhis_dups)
  readr::write_csv(
    df_ncvhis_dups,
    paste0(output_dir,"df_ncvhis_dups_",as_of,".csv.gz")
  )
} else {
  print("No duplicates.")
}
str(df_ncvhis_in)
```  

## D. ncvoter  

### Read ncvoter file   

These are the values for status_cd:  

A Active  
D Denied  
I Inactive  
R Removed  
S Temporary  

reason_cd are some finer detail for status_cd  

election_lbl is a 10 char field that seems to hold a date mm/dd/yyyy  
but without any guarantees that it will always be a date.  
For our purposes, treat as a character field and convert to date  
as required.  

### Postgres table def:  

CREATE TABLE public.ncvoter  
(  
    county_id character(3) COLLATE pg_catalog."default" NOT NULL,  
    voter_reg_num character(12) COLLATE pg_catalog."default" NOT NULL,  
    race_code character(3) COLLATE pg_catalog."default" NOT NULL,  
    ethnic_code character(3) COLLATE pg_catalog."default" NOT NULL,  
    gender_code character(1) COLLATE pg_catalog."default",  
    birth_year integer,  
    status_cd character(1) COLLATE pg_catalog."default" NOT NULL,  
    reason_cd character(2) COLLATE pg_catalog."default",  
    as_of_date date NOT NULL,  
    registr_dt date,  
    CONSTRAINT ncvoter_pkey PRIMARY KEY (county_id, voter_reg_num)  
)  
WITH (  
    OIDS = FALSE  
)  
TABLESPACE pg_default;  
  
ALTER TABLE public.ncvoter  
    OWNER to postgres;  
COMMENT ON TABLE public.ncvoter  
    IS 'From NCSBE ncvoter data';  

### Handling the data file ncvoter_Statewide.txt    
Files are located in data/ncvoter/  
Use ncvoter_Statewide.txt  
UNZIP ncvoter_Statewide.zip prior to this script  
Easier than repeatedly unzipping  

FILE COLS:  
"county_id"	"county_desc"	"voter_reg_num"	"status_cd"	"voter_status_desc"	"reason_cd"	"voter_status_reason_desc"	"absent_ind"	"name_prefx_cd"	"last_name"	 
"first_name"	"middle_name"	"name_suffix_lbl"	"res_street_address"	"res_city_desc"	"state_cd"	"zip_code"	"mail_addr1"	"mail_addr2"	"mail_addr3"	
"mail_addr4"	"mail_city"	"mail_state"	"mail_zipcode"	"full_phone_number"	"race_code"	"ethnic_code"	"party_cd"	"gender_code"	"birth_age"	 
"birth_state"	"drivers_lic"	"registr_dt"	"precinct_abbrv"	"precinct_desc"	"municipality_abbrv"	"municipality_desc"	"ward_abbrv"	"ward_desc"	"cong_dist_abbrv"	 
"super_court_abbrv"	"judic_dist_abbrv"	"nc_senate_abbrv"	"nc_house_abbrv"	"county_commiss_abbrv"	"county_commiss_desc"	"township_abbrv"	"township_desc"	"school_dist_abbrv"	"school_dist_desc"	
"fire_dist_abbrv"	"fire_dist_desc"	"water_dist_abbrv"	"water_dist_desc"	"sewer_dist_abbrv"	"sewer_dist_desc"	"sanit_dist_abbrv"	"sanit_dist_desc"	"rescue_dist_abbrv"	"rescue_dist_desc"	
"munic_dist_abbrv"	"munic_dist_desc"	"dist_1_abbrv"	"dist_1_desc"	"dist_2_abbrv"	"dist_2_desc"	"confidential_ind"	"birth_year"	"ncid"	"vtd_abbrv"	"vtd_desc"  

This is how to unzip from a script if desired (using the utility 7-Zip):  
system(paste0('"c:/program files/7-zip/7z.exe" x ', data_dir ,files, ' -o', data_dir,' -y -r'))  

8/22/18 add ncid  

```{r gray}
#
# HEADER IS IN FILE SINCE THERE WAS NO GREPPING
#
# keep these columns (drops precinct among others):
#
file_cols_keep <- c(
  "county_id",
  "voter_reg_num",
  "status_cd",
  "voter_status_desc",
  "reason_cd",
  "voter_status_reason_desc",
  "party_cd",
  "race_code",
  "ethnic_code",
  "gender_code",
  "registr_dt",
  "birth_year",
  "ncid"
)
#
# fast read using data.table
# convert to data frame for ease of computation
#
df_ncvoter_in <- data.table::fread(
  paste0(data_ncvoter_dir,ncvoter_file),
  select=file_cols_keep
) %>%
as.data.frame()
```  

Clarify codes and keep on hand  
```{r bitter}
df_ncvoter_codes_dcast <- 
  df_ncvoter_in %>% 
  reshape2::dcast(
    race_code+gender_code+ethnic_code~.
  ) %>%
  plyr::rename(
    c("."="N")
  )
#
df_ncvoter_reason_dcast <- 
  df_ncvoter_in %>% 
  reshape2::dcast(
    status_cd+voter_status_desc+reason_cd+voter_status_reason_desc~.
  ) %>%
  plyr::rename(
    c("."="N")
  )
#
```  
Heal ncvoter data frame  

IMPORTANT  
Inspection of the above dcasts shows some missing data  
Replace as appropriate  

```{r hoyt}
df_ncvoter_in %<>%
dplyr::select(
  -voter_status_desc,
  -voter_status_reason_desc
) %>%
dplyr::mutate_if(
  is.character, funs(ifelse(trimws(.)=="",NA,.)) 
)  %>%
dplyr::mutate(
  gender_code=ifelse(is.na(gender_code),"U",gender_code),
  race_code=ifelse(is.na(race_code),"U",race_code)
)
``` 

Coerce to dates  

```{r revel}
df_ncvoter_in %<>%
dplyr::mutate(
  registr_dt=lubridate::as_date(lubridate::mdy(registr_dt)),
  as_of_date=lubridate::as_date(as_of)
)
```  

```{r sybil}
str(df_ncvoter_in)
```

Counts by status_cd:   

```{r ruby}
(df_ncvoter_in_status_cd_dcast <-
df_ncvoter_in %>% 
  reshape2::dcast(
    status_cd~.,
    value.var="as_of_date",
    margins=TRUE
  ) %>%
  plyr::rename(
    c("."="N")
  )
)
#
persons_df_ncvoter_in_all <- nrow(df_ncvoter_in)
```  

## E. df_nc_voted

### Data frame of those who voted  

```{r quail}
df_nc_voted <- 
  dplyr::inner_join(
    df_ncvoter_in %>% 
      dplyr::select(
        -registr_dt
    ),
    df_ncvhis_in %>% 
      dplyr::select(
        county_id,
        voter_reg_num
    ),
    by=c("county_id","voter_reg_num")
  )
```  

```{r waxen}
str(df_nc_voted)
```  

This shows that `r nrow(df_nc_voted)` persons voted.  
The status_cd count shows that about 10% of the Inactive voters (as per df_nc_voter_in

```{r yoke}
(df_ncvoted_status_dcast <- 
  df_nc_voted %>% 
  reshape2::dcast(
    status_cd~.,
    value.var="as_of_date",
    margins=TRUE
  ) %>%
  plyr::rename(
    c("."="N")
  )
)
```  

## F. Finish  

Save large data frames to pare down size of image  

```{r equip}
save(df_ncvoter_in,file="df_ncvoter_in.RData")
save(df_ncvhis_in,file="df_ncvhis_in.RData")
save(df_nc_voted,file="df_nc_voted.RData")
save(df_regis_stats_20161108,file="df_regis_stats_20161108.RData")
```  

```{r tiger}
readr::write_csv(
  df_ncvoter_in,
  paste0(output_dir,"df_ncvoter_in_",as_of,".csv.gz")
)
```  

```{r fuel}
readr::write_csv(
  df_ncvhis_in,
  paste0(output_dir,"df_ncvhis_in_",as_of,".csv.gz")
)
```  

```{r mash}
readr::write_csv(
  df_nc_voted,
  paste0(output_dir,"df_nc_voted_",as_of,".csv.gz")
)
```  

```{r medal}
rm(
  df_ncvoter_in,
  df_ncvhis_in,
  df_nc_voted,
  df_regis_stats_20161108
)
```  

```{r arrow}
save.image("AWR_NCSBE.v2.RData")
```  

```{r lucia}
ls()
```  

## END