---
title: "AreWeRepresented Census and NCACC Categorical Data V2"
output:  
  html_document: default
author: "D. Hopp"
date: "May 14, 2018"
---

INPUT:  
AWR_Census_NCACC.RData  
df_nc_voted.RData  

The AWR_Census_NCACC.RData image includes the directory paths data_dir and output_dir  

OUTPUT:  
df_merge_county_sex_race_20161108.csv.gz  
df_cc_est2016_compare.csv.gz  
df_cc_est2016_compare_Hisp.csv.gz
df_merge_county_sex_race_OMU_20161108.csv.gz - this is a view, see v_df_merge_county_sex_race_OMU_20161108()  
AWR_Census_NCACC_cat.RData an amalgamation of AWR_Census_NCACC.RData and objects created in this script  

## A. Who is Using Which Codes for What  

```{r garb}
load("AWR_NCSBE.v2.RData")
load("df_nc_voted.RData")
load("df_ncvoter_in.RData")
load("df_regis_stats_20161108.RData")
```  

## B. NCSBE  
### Registration Data Race and Ethnicity   

results_pct_20161108.txt via  
https://dl.ncsbe.gov/index.html?prefix=ENRS/2016_11_08/    

Codelists:  

```{r rsvp}  
dfcl_regis_party_cd
dfcl_regis_race_code 
dfcl_regis_ethnic_code
dfcl_regis_sex_code
```  
Convenience data frame:  
Race (only) by county in wide format:  

```{r myth}
fn_df_regis_stats_20161108_county_race_wide <- function() {
  df_ <-
  df_regis_stats_20161108 %>%
  dplyr::select(
    total_voters,
    race_code,
    FIPS3
  ) %>%
  reshape2::dcast(
    FIPS3~race_code,
    value.var = "total_voters",
    sum
  )
  #
  return(df_)
}
```  

Convenience data frame:  
Race (only) by county in long format:  

```{r froth}
fn_df_regis_stats_20161108_county_race_long <- function(){
  df_ <-
  df_regis_stats_20161108 %>%
  dplyr::select(
    total_voters,
    race_code,
    FIPS3
  ) %>%
  dplyr::group_by(
    FIPS3,
    race_code
  ) %>%
  dplyr::summarize(
    N=sum(total_voters)
  ) %>%
  dplyr::ungroup()
  #
  return(df_)
}
```  

```{r calm}
str(fn_df_regis_stats_20161108_county_race_long())
```  

Race and sex by county:  

```{r ward}
fn_df_regis_stats_20161108_county_sex_race <- function(){
  df_ <-
  df_regis_stats_20161108 %>%
  dplyr::select(
    total_voters,
    sex_code,
    race_code,
    FIPS3
  ) %>% 
  dplyr::group_by(
    FIPS3,
    sex_code,
    race_code
  ) %>% 
  dplyr::summarise(
    N=sum(total_voters)
  ) %>%
  dplyr::ungroup()
  #
  return(df_)
}
```  

```{r rhino}
str(fn_df_regis_stats_20161108_county_sex_race())
```  

## C. Census Bureau  
### Race and Ethnicity  

cc-est2016-alldata-37.csv via  
https://www.census.gov/data/tables/2016/demo/popest/counties-detail.html  
"Annual County Resident Population Estimates by Age, Sex, Race, and Hispanic Origin: April 1, 2010 to July 1, 2016"  

Census 2010 Race Methodology: 
[here](https://www2.census.gov/programs-surveys/popest/technical-documentation/methodology/modified-race-summary-file-method/mrsf2010.pdf) 
and [here](https://www.census.gov/prod/cen2010/briefs/c2010br-02.pdf)  

Gender is for each Race, and also total, e.g.,  
TOT_MALE, TOT_FEMALE, WA_MALE, WA_FEMALE etc.  

WA White alone  
BA Black or African American alone  
IA American Indian and Alaska Native alone  
AA Asian alone  
NA Native Hawaiian and Other Pacific Islander alone  
TOM  Two or More Races  

Hispanic is double-counted but included as additional columns:  
H Hispanic  

These are double-counted and are not included:  
WAC  White alone or in combination  
BAC  Black or African American alone  
IAC  American Indian and Alaska Native alone or in combination  

Counts by total population (all) and by voting age (18 and older)  
Breakdown by race, ignoring ethnicity.  
This is for entire state:  

```{r tern}
fn_df_cc_est2016_tot_subset() %>%
dplyr::select(TOT_POP,nWA,nBA,nIA,nAA,nTOM,pop) %>%
reshape2::melt(
  id.vars="pop"
) %>%
reshape2::dcast(
  pop~variable,
  value.var = "value",
  sum
)
```  

This is by county (by FIPS3): df_cc_est2016_county_race_dcast  

```{r mulch}
str(df_cc_est2016_county_race_dcast)
```  

### Reasonable Mappings  

Gender:  
NCSBE sex_code: M, F, U
Census Gender is for each Race, and also total, e.g.,  TOT_MALE, TOT_FEMALE, WA_MALE, WA_FEMALE etc.  

Race:
NCSBE race_code: A, B, I, M, O, U, W (does not include Native Hawaiian explicitly)  
Census Race (see Gender): WA, BA, IA (American Indian), AA, NA (Native Hawaiian etc.), TOM (two or more)  

### Remapping  

Convince census to look like NCSBE data:  

```{r maria}
fn_df_cc_est2016_county_sex_race_long_recoded <- function(){
  #
  #lst_colHead_code <- list(
  #  "WA_"="W","BA_"="B","IA_"="I","AA_"="A","TOM_"="M"
  #)
  #lst_colHead_code_names <- names(lst_colHead_code)
  recode_colHead <- "'WA_MALE'='W'; 'BA_MALE'='B'; 'IA_MALE'='I'; 'AA_MALE'='A'; 'TOM_MALE'='M'; 'WA_FEMALE'='W'; 'BA_FEMALE'='B'; 'IA_FEMALE'='I'; 'AA_FEMALE'='A'; 'TOM_FEMALE'='M'; else='O'"
  #
  df_ <-
  df_cc_est2016_tot %>%
  dplyr::select( 
    FIPS3,
    WA_MALE,
    WA_FEMALE,
    BA_MALE,
    BA_FEMALE,
    IA_MALE,
    IA_FEMALE,
    AA_MALE,
    AA_FEMALE,
    NA_MALE,
    NA_FEMALE,
    TOM_MALE,
    TOM_FEMALE,
    pop
  ) %>%
  reshape2::melt(
    id.vars=c("pop","FIPS3")
  ) %>%
  dplyr::mutate(
    variable=as.character(variable)
  ) %>%
  dplyr::mutate(
    sex=
      ifelse(
        stringr::str_detect(variable,"FEMALE"),"F","M"
      ),
    race=
      car::recode(variable,recode_colHead)
  ) %>%
  dplyr::select(
    pop,
    FIPS3,
    sex,
    race,
    value
  ) %>%
  dplyr::rename(
    N=value
  ) %>%
  dplyr::arrange(
    pop,
    FIPS3,
    sex,
    race
  )
  #
  return(df_)
}
```   

```{r abba}
str(fn_df_cc_est2016_county_sex_race_long_recoded())
df_cc_est2016_county_sex_race_long_recoded <- fn_df_cc_est2016_county_sex_race_long_recoded()
```  

## D. NCACC    

Notice that categories do NOT provide sex vs. race, just totals.  
In that sense, NCACC data is not compatible with census or NCSBE.  
Keep in mind that race most likely is self-reported.  

```{r churn}
fn_df_NCACC_2016_demog_subset <- function(){
  df_ <-
  df_NCACC_2016_demog %>%
  dplyr::mutate(
    nmale=ncomm-nfemale,
    nother_race=ncomm-(nwhite+nafam+namin)
  ) %>%
  dplyr::select(
    FIPS3,
    nmale,
    nfemale,
    nwhite,
    nafam,
    namin,
    nother_race
  )
  #
  return(df_)
}
```  

```{r polka}
str(fn_df_NCACC_2016_demog_subset())
```  

## NCSBE Voting Data  

```{r burma}
df_nc_voted_subset_long <-
df_nc_voted %>%
dplyr::filter(
  status_cd=="A"
) %>%
dplyr::inner_join(
  df_FIPS %>%
    dplyr::select(
      county_id,
      FIPS3
    ),
  by="county_id"
) %>%
dplyr::select(
  FIPS3,
  gender_code,
  race_code
) %>%
reshape2::dcast(
  FIPS3+gender_code+race_code~.
) %>%
plyr::rename(
  c(
    "."="N",
    "race_code"="race",
    "gender_code"="sex"
  )
) %>%
na.omit() %>%
dplyr::arrange(
  FIPS3,
  sex,
  race
)
```  

```{r yodel}
str(df_nc_voted_subset_long)
```  

## E. Bring Together  

Create one data frame to hold all of above (long shape)  
* Columns:  
* pop: 
  + all: Census Bureau all population  
  + voteage: Census Bureau 18 years and older  
  + regis: NCSBE registered to vote  
  + voted: NCSBE voted  
* sex:  
  + F: Female  
  + M: Male  
  + U: Unknown  
* race:  
  A: Asian  
  B: African-American  
  I: American Indian  
  M: More than one  
  O: Other  
  U: Unknown  
  W: White  

```{r groton}
df_merge_county_sex_race_20161108 <-
  rbind(
    fn_df_regis_stats_20161108_county_sex_race() %>%
    dplyr::rename(
      sex=sex_code,
      race=race_code
    ) %>%
    dplyr::mutate(
      pop="regis"
    ) %>%
    dplyr::select(
      pop,
      FIPS3,
      sex,
      race,
      N
    ),
    df_cc_est2016_county_sex_race_long_recoded %>%
      dplyr::filter(
        race != "all"
      ),
    df_nc_voted_subset_long %>%
    dplyr::mutate(
      pop="voted"
    ) %>%
    dplyr::select(
      pop,
      FIPS3,
      sex,
      race,
      N
    )
  )
```  

```{r icy}
str(df_merge_county_sex_race_20161108)
```  

Write df_merge_county_sex_race_20161108 as a csv file.  

```{r pencil}
readr::write_csv(
  df_merge_county_sex_race_20161108,
  paste0(output_dir,"df_merge_county_sex_race_20161108.csv.gz")
)
```  

Any analysis has to be aware of the numbers by race, which present comparison challenges across the sources.  
The "O" and "U" categories are of particular note.  

O (Other) and U (Unknown) are used in the NCSBE data, as per ncvhis_ncvoter_data_format.txt but only O in the Census data.  
Are both O designations to be considered comparable?  

Display counts by race:  

```{r spire}
fn_merge_county_sex_race_20161108_dcast <- function(){
  # all this to get a nice lookin dcast
  # there must be a better way
  # use an ordered factor
  #
  df_ <- df_merge_county_sex_race_20161108 
  df_$pop <- as.factor(df_$pop)
  df_$pop <- ordered(
    df_$pop, 
    levels=c("all","voteage","regis","voted")
  )
  #
  df_dcast <-
  df_ %>% 
  dplyr::select(
    pop,
    race,
    N
  ) %>% 
  reshape2::dcast(
    race~pop, 
    value.var="N",
    sum,  
    margins="race"
  )
  #
  return(
    df_dcast
  )
}
```    

```{r etch}
fn_merge_county_sex_race_20161108_dcast()
```  

### This combines the df_merge_county_sex_race_20161108 race categories M, O, and U into O alone.  

```{r evade}
v_df_merge_county_sex_race_OMU_20161108 <- function(){
  df_ <-
  df_merge_county_sex_race_20161108 %>%
    dplyr::mutate(
      race=ifelse(race %in% c("M","U"),"O",race)
    ) %>%
  reshape2::dcast(
    pop+FIPS3+sex+race~.,
    value.var="N",
    sum
  ) %>%
  plyr::rename(
    c("."="N")
  )
  #
  return(df_)
}
```

```{r garner}
str(v_df_merge_county_sex_race_OMU_20161108())
```  

Check on counts:

```{r hunt}
df_merge_county_sex_race_20161108 %>%
  dplyr::filter(
    race %in% c("O","M","U")
  ) %>%
  reshape2::dcast(
    pop+race~.,
    value.var="N",
    sum
  )
```  

```{r prep}
v_df_merge_county_sex_race_OMU_20161108() %>%
  dplyr::filter(
    race == "O"
  ) %>%
  reshape2::dcast(
    pop+race~.,
    value.var="N",
    sum
  )
```

```{r widen}
readr::write_csv(
  v_df_merge_county_sex_race_OMU_20161108(),
  paste0(output_dir,"df_merge_county_sex_race_OMU_20161108.csv.gz")
)
```  
## F. How do the Census Bureau race classification sets compare?  

### Discussion  

These are the column headings in cc-est2016-alldata-37.csv as described in cc-est2016-alldata.pdf  

SUMLEV,STATE,COUNTY,STNAME,CTYNAME,YEAR,AGEGRP,  
TOT_POP,TOT_MALE,TOT_FEMALE,  
WA_MALE,WA_FEMALE,BA_MALE,BA_FEMALE,IA_MALE,IA_FEMALE,AA_MALE,AA_FEMALE,NA_MALE,NA_FEMALE,TOM_MALE,TOM_FEMALE,  
WAC_MALE,WAC_FEMALE,BAC_MALE,BAC_FEMALE,IAC_MALE,IAC_FEMALE,AAC_MALE,AAC_FEMALE,NAC_MALE,NAC_FEMALE,  
NH_MALE,NH_FEMALE,NHWA_MALE,NHWA_FEMALE,NHBA_MALE,NHBA_FEMALE,NHIA_MALE,NHIA_FEMALE,NHAA_MALE,NHAA_FEMALE,
NHNA_MALE,NHNA_FEMALE,NHTOM_MALE,NHTOM_FEMALE,
NHWAC_MALE,NHWAC_FEMALE,NHBAC_MALE,NHBAC_FEMALE,NHIAC_MALE,NHIAC_FEMALE,NHAAC_MALE,NHAAC_FEMALE,NHNAC_MALE,NHNAC_FEMALE,
H_MALE,H_FEMALE,HWA_MALE,HWA_FEMALE,HBA_MALE,HBA_FEMALE,HIA_MALE,HIA_FEMALE,HAA_MALE,HAA_FEMALE,HNA_MALE,
HNA_FEMALE,HTOM_MALE,HTOM_FEMALE,
HWAC_MALE,HWAC_FEMALE,HBAC_MALE,HBAC_FEMALE,HIAC_MALE,HIAC_FEMALE,HAAC_MALE,HAAC_FEMALE,HNAC_MALE,HNAC_FEMALE  

Question: Are there any significant differences in counts when looking at the "WA_MALE..." and the "WAC_MALE..." categorizations? For now, we will ignore the Hispanic categorizations (NH_MALE...). Notice that there is no Unknown category. We will refer to the WA_... as Source A, and the WAC_... as Source B.  

The following is from https://www2.census.gov/programs-surveys/popest/technical-documentation/methodology/2010-2016/2016-natstcopr-meth.pdf   

 "we modify the 2010 Census race categories to be consistent with the race categories that appear in our input data. The 2010 Census allowed for responses that included one or more of the race groups defined by the Office of Management and Budget (OMB) in 1997: White; Black or African American; American Indian and Alaska Native; Asian; and Native Hawaiian and Other Pacific Islander. The 2010 Census also allowed responses that included “Some other race.” The Population Estimates program produces estimates in five race categories and their combinations. "Some other race" responses have been “modified” to be consistent with these race
categories.  

This modification process recategorizes responses that include “Some other race” alone or in combination with other races into one or more of the five OMB race categories. When a “Some other race” response appears in combination with one or more of the OMB races, we remove the “Some other race” category. Responses that
include “Some other race” alone are allocated to one or more of the five OMB categories listed above using information on the household when available or a hot decking procedure if necessary.  

Note that this editing process produces tabulations for our estimates that show fewer people reporting two or more races than similar tabulations from the 2010 Census. This is primarily because responses of “Some other race” and one of the OMB defined races in the 2010 Census appear in the single OMB race category in the
estimates base."  

### Conclusion  

This conclusion is substantiated in the script that follows below. It is presented here for clarity.  

The problem with these talleys seems to be that the "HWA_..." (including HTOM_) sum equals H_MALE+H_FEMALE,  
while the "HWAC_..." series sum is larger. Whatever the source of this, for our purposes the Hispanic counts  
will be represented by the "HWA_..." series.  

## G. Create R Objects  

Utility data frame

```{r peel}
df_cc_est2016_tot_only <-
  df_cc_est2016_tot %>%
  dplyr::select(
    pop,
    FIPS3,
    TOT_POP,
    TOT_MALE,
    TOT_FEMALE
  )
```  

```{r lydia}
str(df_cc_est2016_tot_only)
```  

Function to create a data frame used in comparison of the two categorizations as per above.  

```{r round}  
fn_df_cc_est2016_compare <- function(){
  #
  # Use vectors to parse column headings.
  #
  # 4 categories + Other (which includes TOM_)
  #
  #colHead_A_code <- c("WA_","BA_","IA_","AA_")
  #colHead_A_abbrev <- c("W","B","I","A")
  #lst_colHead_A_code <- list(
  #  "WA_"="W","BA_"="B","IA_"="I","AA_"="A"
  #)
  #lst_colHead_A_code_names <- names(lst_colHead_A_code)
  recode_colHead_A <- 
    "'WA_MALE'='W'; 'BA_MALE'='B'; 'IA_MALE'='I'; 'AA_MALE'='A'; 'WA_FEMALE'='W'; 'BA_FEMALE'='B'; 'IA_FEMALE'='I'; 'AA_FEMALE'='A'; else='O'"
  #
  df_ <-
  df_cc_est2016_tot %>%
  dplyr::select( 
    FIPS3,
    WA_MALE,
    WA_FEMALE,
    BA_MALE,
    BA_FEMALE,
    IA_MALE,
    IA_FEMALE,
    AA_MALE,
    AA_FEMALE,
    NA_MALE,
    NA_FEMALE,
    TOM_MALE,
    TOM_FEMALE,
    pop
    ) %>%
  reshape2::melt(
    id.vars=c("pop","FIPS3")
  ) %>%
  dplyr::mutate(
    variable=as.character(variable)
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    sex=
      ifelse(
        stringr::str_detect(variable,"FEMALE"),
        "F",
        "M"
      ),
  race=
    car::recode(variable,recode_colHead_A),
  source="A"
  ) %>%
  data.frame() %>%
  dplyr::select(
    pop,
    source,
    FIPS3,
    sex,
    race,
    value
  ) 
  #
  # 4 categories + Other
  #
  #lst_colHead_B_code <- list(
  #  "WAC_"="W","BAC_"="B","IAC_"="I","AAC_"="A"
  #)
  #lst_colHead_B_code_names <- names(lst_colHead_B_code)
  recode_colHead_B <- 
    "'WAC_MALE'='W'; 'BAC_MALE'='B'; 'IAC_MALE'='I'; 'AAC_MALE'='A'; 'WAC_FEMALE'='W'; 'BAC_FEMALE'='B'; 'IAC_FEMALE'='I'; 'AAC_FEMALE'='A'; else='O'"
  #
  df_B <-
  df_cc_est2016_tot %>%
  dplyr::select( 
    FIPS3,
    WAC_MALE,
    WAC_FEMALE,
    BAC_MALE,
    BAC_FEMALE,
    IAC_MALE,
    IAC_FEMALE,
    AAC_MALE,
    AAC_FEMALE,
    NAC_MALE,
    NAC_FEMALE,
    pop
    ) %>%
  reshape2::melt(
    id.vars=c("pop","FIPS3")
  ) %>%
  dplyr::mutate(
    variable=as.character(variable)
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    sex=
      ifelse(
        stringr::str_detect(variable,"FEMALE"),
        "F",
        "M"
      ),
  race=
    car::recode(variable,recode_colHead_B),
    source="B"
  ) %>%
  data.frame() %>%
  dplyr::select(
    pop,
    source,
    FIPS3,
    sex,
    race,
    value
  ) 
  # 
  # summarize
  #
  df_ <-
  rbind(
    df_,
    df_B
  ) %>%
  dplyr::group_by(
    pop,
    source,
    FIPS3,
    sex,
    race
  ) %>%
  dplyr::summarize(
    N=sum(value)
  ) %>%
  dplyr::ungroup()
  #
  return(df_)
}
```  

```{r briar}
df_cc_est2016_compare <- fn_df_cc_est2016_compare()
str(df_cc_est2016_compare) 
```  
Write as a csv file

```{r oat}
readr::write_csv(
  df_cc_est2016_compare,
  paste0(output_dir,"df_cc_est2016_compare.csv.gz")
)
```

View to sum male and female  

```{r baton}
v_df_cc_est2016_compare <- function(){
  fn_df_cc_est2016_compare() %>%
  dplyr::group_by(
    pop,
    source,
    FIPS3,
    race
  ) %>%
  dplyr::summarize(
    N_F_M=sum(N)
  ) %>%
  dplyr::ungroup()
}
```  

```{r tunnel}
str(v_df_cc_est2016_compare())
```  

### Compare source A vs B to try to see what race count differences might be.  

```{r stamp}
# drop sex
df_cc_est2016_compare_N <- v_df_cc_est2016_compare()
#
df_cc_est2016_compare_N_source <-
dplyr::inner_join(
  df_cc_est2016_compare_N %>%
    dplyr::filter(
      source=="A" &
      pop=="all"
    ),
  df_cc_est2016_compare_N %>%
    dplyr::filter(
      source=="B" &
      pop=="all"
    ),  
  by=c("pop","FIPS3","race")
) %>%
dplyr::select(
  pop,
  FIPS3,
  race,
  N_F_M.x,
  N_F_M.y
) %>%
dplyr::rename(
  N.A=N_F_M.x,
  N.B=N_F_M.y
) %>%
dplyr::mutate(
  pctN.A=round(N.A/(N.A+N.B),digits=3),
  logitN.A=round(log((N.A/(N.A+N.B))/(1-(N.A/(N.A+N.B)))),digits=3)
)
```  

```{r mabel}
str(df_cc_est2016_compare_N_source)
```  

### Quintiles  

```{r bulb}
fn_df_cc_est2016_compare_N_quintile <- function(race_){
  ret <- list()
  df_ <-
  df_cc_est2016_compare_N_source %>%
    dplyr::filter(
      race==race_
    ) 
  ret[[1]] <- median(df_[["logitN.A"]])
  ret[[2]] <- quantile(df_[["logitN.A"]],probs=seq(0,1,.2))
  #
  return(ret)
}
#
B_df_cc_est2016_compare_N_quintile <- fn_df_cc_est2016_compare_N_quintile("B")
W_df_cc_est2016_compare_N_quintile <- fn_df_cc_est2016_compare_N_quintile("W")
```  

The Black median is `r B_df_cc_est2016_compare_N_quintile[[1]]` and the White median is `r W_df_cc_est2016_compare_N_quintile[[1]]`.  

The quintiles for Blacks are  
```{r sears}
B_df_cc_est2016_compare_N_quintile[[2]]
```  
and for Whites are  
```{r rebut}
W_df_cc_est2016_compare_N_quintile[[2]]
```  

### Cleveland Plots  

Here are Cleveland plots for the logit of Source A (WA_...) proportion. A logit of 0 means that N.A=N.B, since the total is N.A+N.B. Negative values mean that N.A<N.B. Logit used, rather than percentage, since there is a great range in total count over the counties. Reporting percentages would be associated with larger erros for the smaller counties.  

The two plots show a remarkable difference in scale, the difference between N.A and N.B being much larger for Black than for White.  

White  

```{r rest, fig.width=10,fig.height=12}
df_cc_est2016_compare_N_source %>%
  dplyr::filter(
    race=="W"
  ) %>%
  ggplot( 
    aes(y=reorder(FIPS3,logitN.A), x=logitN.A) 
  ) +
  geom_point(size=3) +
  scale_x_continuous(breaks=seq(-.1,0,.01)) +
  labs(
    title="Logit for Whites Comparing Source A to Source B",
    subtitle="log((N.A/(N.A+N.B))/(1-(N.A/(N.A+N.B))))"
  ) +
  ylab("County FIPS3") +
  theme_bw() +
  theme(
    plot.title=element_text(hjust=0.5),
    plot.subtitle=element_text(hjust=0.5),
    panel.grid.major.x=element_blank(),
    panel.grid.minor.x=element_blank(),
    panel.grid.major.y=element_line(colour="grey60",linetype="dashed")
  )
```  

Black  

```{r dial, fig.width=10, fig.height=12}
df_cc_est2016_compare_N_source %>%
  dplyr::filter(
    race=="B"
  ) %>%
  ggplot( 
    aes(y=reorder(FIPS3,logitN.A), x=logitN.A) 
  ) +
  geom_point(size=3) +
  scale_x_continuous(breaks=seq(-1.0,0,.1)) +
  labs(
    title="Logit for Blacks Comparing Source A to Source B",
    subtitle="log((N.A/(N.A+N.B))/(1-(N.A/(N.A+N.B))))"
  ) +
  ylab("County FIPS3") +
  theme_bw() +
  theme(
    plot.title=element_text(hjust=0.5),
    plot.subtitle=element_text(hjust=0.5),
    panel.grid.major.x=element_blank(),
    panel.grid.minor.x=element_blank(),
    panel.grid.major.y=element_line(colour="grey60",linetype="dashed")
  )
```  

## H. Considering the Hispanic Census Bureau Categories  

```{r crab}
fn_df_cc_est2016_compare_Hisp <- function(){
  #
  # Use vectors to parse column headings.
  #  
  #lst_colHead_HA_code <- list(
  #  "NHWA_"="W","HWA_"="W","NHBA_"="B","HBA_"="B","NHIA_"="I","HIA_"="I","NHAA_"="A","HAA_"="A" 
  #)
  #lst_colHead_HA_code_names <- names(lst_colHead_HA_code)
  recode_colHeadHA <- "'NHWA_MALE'='W'; 'HWA_MALE'='W'; 'NHBA_MALE'='B'; 'HBA_MALE'='B'; 'NHIA_MALE'='I'; 'HIA_MALE'='I'; 'NHAA_MALE'='A'; 'HAA_MALE'='A'; 'NHWA_FEMALE'='W'; 'HWA_FEMALE'='W'; 'NHBA_FEMALE'='B'; 'HBA_FEMALE'='B'; 'NHIA_FEMALE'='I'; 'HIA_FEMALE'='I'; 'NHAA_FEMALE'='A'; 'HAA_FEMALE'='A'; else='O'"
  #
  df_ <-
  df_cc_est2016_tot %>%
  dplyr::select( 
    FIPS3,
    NHWA_MALE,
    NHWA_FEMALE,
    NHBA_MALE,
    NHBA_FEMALE,
    NHIA_MALE,
    NHIA_FEMALE,
    NHAA_MALE,
    NHAA_FEMALE,
    NHNA_MALE,
    NHNA_FEMALE,
    NHTOM_MALE,
    NHTOM_FEMALE,
    HWA_MALE,
    HWA_FEMALE,
    HBA_MALE,
    HBA_FEMALE,
    HIA_MALE,
    HIA_FEMALE,
    HAA_MALE,
    HAA_FEMALE,
    HNA_MALE,
    HNA_FEMALE,
    HTOM_MALE,
    HTOM_FEMALE,
    pop
    ) %>%
  reshape2::melt(
    id.vars=c("pop","FIPS3")
  ) %>%
  dplyr::mutate(
    variable=as.character(variable)
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    sex=
      ifelse(
        stringr::str_detect(variable,"FEMALE"),
        "F",
        "M"
      ),
    race=
      car::recode(variable,recode_colHeadHA),
    hisp=
      ifelse(
        stringr::str_sub(variable,1,1)=="H",
        TRUE,
        FALSE
      ),
    source="A"
  ) %>%
  data.frame() %>%
  dplyr::select(
    pop,
    source,
    FIPS3,
    sex,
    race,
    hisp,
    value
  )
  #
  #lst_colHead_HB_code <- list(
  #  "NHWAC_"="W","HWAC_"="W","NHBAC_"="B","HBAC_"="B","NHIAC_"="I","HIAC_"="I","NHAAC_"="A","HAAC_"="A" 
  #)
  #lst_colHead_HB_code_names <- names(lst_colHead_HB_code)
  recode_colHeadHB <- "'NHWAC_MALE'='W'; 'HWAC_MALE'='W'; 'NHBAC_MALE'='B'; 'HBAC_MALE'='B'; 'NHIAC_MALE'='I'; 'HIAC_MALE'='I'; 'NHAAC_MALE'='A'; 'HAAC_MALE'='A'; 'NHWAC_FEMALE'='W'; 'HWAC_FEMALE'='W'; 'NHBAC_FEMALE'='B'; 'HBAC_FEMALE'='B'; 'NHIAC_FEMALE'='I'; 'HIAC_FEMALE'='I'; 'NHAAC_FEMALE'='A'; 'HAAC_FEMALE'='A'; else='O'"
  #
  df_B <-
  df_cc_est2016_tot %>%
  dplyr::select( 
    FIPS3,
    NHWAC_MALE,
    NHWAC_FEMALE,
    NHBAC_MALE,
    NHBAC_FEMALE,
    NHIAC_MALE,
    NHIAC_FEMALE,
    NHAAC_MALE,
    NHAAC_FEMALE,
    NHNAC_MALE,
    NHNAC_FEMALE,
    HWAC_MALE,
    HWAC_FEMALE,
    HBAC_MALE,
    HBAC_FEMALE,
    HIAC_MALE,
    HIAC_FEMALE,
    HAAC_MALE,
    HAAC_FEMALE,
    HNAC_MALE,
    HNAC_FEMALE,
    pop
    ) %>%
  reshape2::melt(
    id.vars=c("pop","FIPS3")
  ) %>%
  dplyr::mutate(
    variable=as.character(variable)
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    sex=
      ifelse(
        stringr::str_detect(variable,"FEMALE"),
        "F",
        "M"
      ),
    race=
      car::recode(variable,recode_colHeadHB),
    hisp=
      ifelse(
        stringr::str_sub(variable,1,1)=="H",
        TRUE,
        FALSE
      ),
    source="B"
  ) %>%
  data.frame() %>%
  dplyr::select(
    pop,
    source,
    FIPS3,
    sex,
    race,
    hisp,
    value
  ) 
  # 
  # summarize
  #
  df_ <-
  rbind(
    df_,
    df_B
  ) %>%
  dplyr::group_by(
    pop,
    source,
    FIPS3,
    sex,
    race,
    hisp
  ) %>%
  dplyr::summarize(
    N=sum(value)
  ) %>%
  dplyr::ungroup()
  #
  return(df_)  
}
```  

```{r zurich}
df_cc_est2016_compare_Hisp <- fn_df_cc_est2016_compare_Hisp()
str(df_cc_est2016_compare_Hisp)
```  

Write as a csv file

```{r pride}
readr::write_csv(
  df_cc_est2016_compare_Hisp,
  paste0(output_dir,"df_cc_est2016_compare_Hisp.csv.gz")
)
```  

### Problem in Counts  

The data frames generated above share in a problem illustrated by these two tables:  

Data not broken down into Hispanic:
```{r oar}
df_cc_est2016_compare %>% 
  reshape2::dcast(pop+source~.,value.var = "N",sum)
```  

Data broken down into Hispanic:  
```{r giant}
df_cc_est2016_compare_Hisp %>% 
  reshape2::dcast(pop+source~.,value.var = "N",sum)
```  

There are mismatches between the source A and B data for the total counts. The Source A figure for total population agrees with that explicitly in the Census Bureau cc_est2016 file. Source B (the WAC... etc.) are the larger, so it appears that there may be a double counting somewhere in Source B.  

The sum for Source A for data not broken down into Hispanic (the first of the dcasts) matches the sum from df_cc_est2016_tot.  
```{r trawl}
df_cc_est2016_tot %>% reshape2::dcast(pop~.,value.var = "TOT_POP",sum)
```  

This matches the sum from df_cc_est2016_alldata_37  
```{r lush}
sum(df_cc_est2016_alldata_37 %>% 
  dplyr::filter(YEAR==6 & AGEGRP==0) %>% 
  dplyr::select(TOT_POP)
)
```  

Looking at totals:

```{r}
df_cc_est2016_alldata_37 %>% 
  dplyr::filter(YEAR==6 & AGEGRP==0) %>% 
  dplyr::select(
    TOT_POP,
    TOT_MALE,
    TOT_FEMALE,
    NH_MALE,
    NH_FEMALE,
    H_MALE,
    H_FEMALE
  ) %>% 
  apply(2,sum)
```  

and in more detail:

```{r excel}
count_cc_est2016_alldata_37 <-
df_cc_est2016_alldata_37 %>% 
  dplyr::filter(YEAR==6 & AGEGRP==0) %>% 
  dplyr::select(
    -SUMLEV,
    -STATE,
    -COUNTY,
    -STNAME,
    -CTYNAME,
    -YEAR,
    -AGEGRP
  ) %>% 
  apply(2,sum)
#
count_cc_est2016_alldata_37
#
df_count_cc_est2016_alldata_37 <-
  data.frame(name=names(count_cc_est2016_alldata_37),N=count_cc_est2016_alldata_37)
```  

The problem with these talleys seems to be that the "HWA_..." (including HTOM_) sum equals H_MALE+H_FEMALE,  
while the "HWAC_..." series sum is larger. Whatever the source of this, for our purposes the Hispanic counts  
will be represented by the "HWA_..." series.  

## Hispanic-based County Counts  

NOTE that df_cc_est2016_compare_Hisp must be filtered to use source "A" only. 

The NCSBE race categories are `r unique(df_merge_county_sex_race_20161108$race)`.  
The Census Bureau race categories are `r t(unique(df_cc_est2016_compare_Hisp %>% dplyr::filter(source=="A") %>% dplyr::select(race)))[1,]`.  
The view of NCSBE v_df_merge_county_sex_race_OMU_20161108() race categories are `r unique(v_df_merge_county_sex_race_OMU_20161108()$race)`.
 
These data frame representations of the data WITHOUT ETHNICITY:  

```{r edge}
df_merge_county_sex_race_20161108 %>% reshape2::dcast(pop+sex+race~.,value.var = "N",sum)
df_cc_est2016_compare_Hisp %>% dplyr::filter(source=="A") %>% reshape2::dcast(pop+sex+race~.,value.var = "N",sum)
v_df_merge_county_sex_race_OMU_20161108() %>% reshape2::dcast(pop+sex+race~.,value.var = "N",sum)
```  

In a more compact display, not including sex WITHOUT ETHNICITY:

```{r spate}
df_merge_county_sex_race_20161108 %>% reshape2::dcast(pop~race,value.var = "N",sum)
df_cc_est2016_compare_Hisp %>% dplyr::filter(source=="A") %>% reshape2::dcast(pop~race,value.var = "N",sum)
v_df_merge_county_sex_race_OMU_20161108() %>% reshape2::dcast(pop~race,value.var = "N",sum)
```  

### The Problem With NCSBE Ethnicity  

Reporting of ethnic_code in df_nc_voted:

```{r onion}
df_nc_voted %>% reshape2::dcast(ethnic_code~.)
```  

The non-responses outnumber the HL responses by a factor of nine. It does not appear reasonable to carry out any analysis of ethnicity of voters.  

## I. COMMENTS ON nc_voter.csv AND voter_stats_20161108.txt  

It is not evident how to compare these two files with confidence. The summary-by-category file, voter_stats_20161108.txt, has `r persons_df_regis_stats_20161108` persons. On the other hand, nc_voter.csv has `r persons_df_ncvoter_in` persons. The difference is `r persons_df_ncvoter_in-persons_df_regis_stats_20161108`. This is close to being accounted for by the sum of categories A (Active) and I (Inactive) in df_ncvoter_in, namely `r df_ncvoter_in_status_cd_dast[1,2] + df_ncvoter_in_status_cd_dast[3,2]`.  


We will be referring to this [NCSBE table](https://vt.ncsbe.gov/RegStat/Results/?date=11%2F08%2F2016)   

This R function uses df_regis_stats_20161108 to attempt to replicate the NCSBE 11/08/2016 table.  
The R code is rather inelegant - any suggestions for improvement would be appreciated.  

```{r illume}
fn_summary_df_regis_stats_20161108 <- function(){
  #
  df1_ <-
  df_regis_stats_20161108 %>%
    dplyr::select(
      FIPS3,
      party_cd,
      total_voters
    ) %>%
    reshape2::dcast(
      FIPS3~party_cd,
      value.var="total_voters",
      sum
    )
  #
  df2_ <-
  df_regis_stats_20161108 %>%
    dplyr::select(
      FIPS3,
      race_code,
      total_voters
    ) %>%
    reshape2::dcast(
      FIPS3~race_code,
      value.var="total_voters",
      sum
    )
  #  
  df3_ <-
  df_regis_stats_20161108 %>%
    dplyr::select(
      FIPS3,
      ethnic_code,
      total_voters
    ) %>%
    reshape2::dcast(
      FIPS3~ethnic_code,
      value.var="total_voters",
      sum
    )
  #  
  df4_ <-
  df_regis_stats_20161108 %>%
    dplyr::select(
      FIPS3,
      sex_code,
      total_voters
    ) %>%
    reshape2::dcast(
      FIPS3~sex_code,
      value.var="total_voters",
      sum
    )
  #  
  df_ <-
  dplyr::inner_join(
    dplyr::inner_join(
      df1_,
      df2_,
      by="FIPS3"
    ),
    dplyr::inner_join(
      df3_,
      df4_,
      by="FIPS3"
    ),
    by="FIPS3"
  ) %>%
  dplyr::rename(
    Asian=A,
    Black=B,
    AmInd=I,
    Multi=M.x,
    Other=O,
    R_UN=U.x,
    White=W,
    E_UN=UN,
    Female=F,
    Male=M.y,
    S_UN=U.y
  ) %>%
  dplyr::mutate(
    Total=Female+Male+S_UN
  )
  #
  return(df_)
}
```  

The results are close to what the NCSBE table shows:  

```{r veal}
fn_summary_df_regis_stats_20161108()
```  

This is the same function as above but tailored to use df_ncvoter_in.  

```{r percy}
fn_summary_df_ncvoter <- function(I=FALSE,D=FALSE,R=FALSE,S=FALSE){
  # for_df_ is df_ncvoter_in subset on status_cd
  # A Active, I Inactive, D Denied, R Removed, S Temporary
  #
  cond_ <- "A"
  if (I) {
    cond_[length(cond_)+1] <- "I"
  }
  if (D) {
    cond_[length(cond_)+1] <- "D"
  }
  if (R) {
    cond_[length(cond_)+1] <- "R"
  }
  if (S) {
    cond_[length(cond_)+1] <- "S"
  }
  for_df_ <- 
  df_ncvoter_in %>%
    dplyr::filter(
      status_cd %in% cond_
    )
  # 
  df1_ <-
  for_df_ %>%
  reshape2::dcast(
    county_id~party_cd
  )
  #
  df2_ <-
  for_df_ %>%
  reshape2::dcast(
    county_id~race_code
  )
  #
  df3_ <-
  for_df_ %>%
  reshape2::dcast(
    county_id~ethnic_code
  )
  #
  df4_ <-
  for_df_ %>%
  reshape2::dcast(
    county_id~gender_code
  ) %>%
  dplyr::rename(
    Female=F,
    Male=M,
    S_UN=U
  )
  # 
  df_ <-
  dplyr::inner_join(
    dplyr::inner_join(
      df1_,
      df2_,
      by="county_id"
    ),
    dplyr::inner_join(
      df3_,
      df4_,
      by="county_id"
    ),
    by="county_id"
  ) %>%
  dplyr::inner_join(
    df_FIPS %>% 
      dplyr::select(
        county_id,
        County
      ),
    by="county_id"
  ) %>%
  dplyr::mutate(
    R_other=A+M+O+U,
    Total=Female+Male+S_UN
  )  %>%
  dplyr::select(
    County,
    DEM,
    REP,
    GRE,
    LIB,
    UNA,
    W,
    B,
    I,
    R_other,
    HL,
    Male,
    Female,
    Total
  )
  #
  return(df_)
}
```  

Including status_cd Active and Inactive gives results close to what the NCSBE table shows:  

```{r rabbit}
fn_summary_df_ncvoter(I=TRUE)
```  

## J. An Important Observation  

As mentioned earlier and throughout this analysis, there is what amounts to a mismatch of the number of Active voters (in df_ncvoter_in) with the number of votes cast (df_ncvhis_in) in the 2016 general election. This is by first considering the status of those voters:

```{r soapy}
df_ncvoted_status_dcast
```  

The above should be compared with the status of all the registered persons:

```{r nonce}
df_ncvoter_in_status_cd_dast
```

Evidently, about 10% of the Inactive persons actually voted.  

## K. Finish  

Do not resave df_nc_voted or df_regis_stats_20161108

```{r moose}
rm(df_nc_voted,df_regis_stats_20161108)
ls()
```

```{r flop}
save.image("AWR_Census_NCACC_cat.RData")
```  

### END  