---
title: "AreWeRepresented Census and NCACC Data Munging V2"
output: 
  html_document: default
author: "D. Hopp"
date: "May 11, 2018"
---

R requires packages plyr, dplyr, reshape2, readr  

INPUT:  
Data files that must be present (see context for sources):  
FIPS.csv  
cc-est2016-alldata-37.csv  
results_pct_20161108.txt  
voter_stats_20161108.txt  
2016 Election Demographics for Website_1_extract.csv  
  
SIGNIFICANT R OBJECTS CREATED:  
df_FIPS  county and FIPS code  
df_cc_est2016_alldata_37  reads df_cc_est2016_alldata_37.csv  
df_cc_est2016_tot  census count of population (includes records for all and for voting age)  
df_results_pct_20161108  reads results_pct_20161108.txt  
df_results_pct_20161108_clean  munged df_results_pct_20161108  
dfcl_results_precinct_Contest.Type.Name  code list  
dfcl_results_precinct_Choice  code list  
df_results_pct_20161108_subset based on df_results_pct_20161108_clean  
df_regis_stats_20161108 reads df_regis_stats_20161108.txt  
dfcl_regis_party_cd code list  
dfcl_regis_race_code code list  
dfcl_regis_ethnic_code  code list  
dfcl_regis_sex_code code list  
df_regis_age_dcast based on df_regis_stats_20161108  
df_regis_stats_20161108_county_rollup county rollup  
df_NCACC_2016_demog read 2016 Election Demographics for Website_1_extract.csv  

OUTPUT:  
df_merge_county_sex_race_20161108.csv.gz
df_NCACC_2016_demog.csv.gz
AWR_Census_NCACC.RData  

DIRECTORY STRUCTURE:  

```{r adieu}
data_dir <- "data/"
output_dir <- "output/"
```  

Check for existence  

```{r natty}
Census_NCACC_files_in_ <- c(
  "FIPS.csv",
  "cc-est2016-alldata-37.csv",
  "results_pct_20161108.txt",
  "voter_stats_20161108.txt",
  "2016 Election Demographics for Website_1_extract.csv"
)
for (f_ in Census_NCACC_files_in_) {
  tmp_ <- paste0(data_dir,f_)
  if (!file.exists(tmp_)) {
    print(paste(tmp_,"does not exist."))
    knitr::knit_exit()
  }
}
#
rm(f_,tmp_)
```  

## A. FIPS Code  

Add 3-char FIPS3 code since some sources use as a numeric, some as character.  
Also add county_id used in NCSBE data.  
This code chunk appears in several places to assure that df_FIPS exists.  

```{r yew}
df_FIPS <- read.csv(
  paste0(data_dir,"FIPS.csv"),
  header=TRUE,
  stringsAsFactors = FALSE
) %>%
dplyr::rowwise() %>%
dplyr::mutate(
   FIPS3=paste0(substr("00",1,3-nchar(as.character(FIPS))),as.character(FIPS))
) %>%
as.data.frame() 
#
df_FIPS %<>%
cbind(
  rownames(df_FIPS)
) %>%
plyr::rename(
  c("rownames(df_FIPS)"="county_id")
) %>%
dplyr::mutate(
  county_id=as.character(county_id)
)
```  

```{r knit}
str(df_FIPS)
```  

## B. County voting-age population counts  
Produce a csv file of county voting-age population counts  

Census 2010 Race Methodology: 
[here](https://www2.census.gov/programs-surveys/popest/technical-documentation/methodology/modified-race-summary-file-method/mrsf2010.pdf) 
and [here](https://www.census.gov/prod/cen2010/briefs/c2010br-02.pdf)  

data source:  
https://www.census.gov/data/tables/2016/demo/popest/counties-detail.html  
to retrieve cc-est2016-alldata-37.csv  
from "Annual County Resident Population Estimates by Age, Sex, Race, and Hispanic Origin: April 1, 2010 to July 1, 2016"  

COUNTY is the FIPS code (watch out for this!)  

we want  
YEAR=6 is 2016      
Note that AGEGRP 0 is a summary figure.  
AGEGRP >= 5 ages 20 and higher  
  note: 4 is ages 15-19  
  take 2/5 of AGEGRP 4 as an estimate of ages 18-19  

Read cc-est2016-alldata-37.csv into data frame  
 
```{r baby}
df_cc_est2016_alldata_37 <- read.csv(  
  paste0(data_dir,"cc-est2016-alldata-37.csv"),
  sep=",",
  header=TRUE,
  stringsAsFactors=FALSE
)
```  

```{r exalt}
str(df_cc_est2016_alldata_37)
```  

Create a function that produces the desired data frame. 
Less clutter this way and intermediate structures can be inspected.  

```{r glad}
fn_df_cc_est2016_tot <- function() {

  # AGEGRP >= 5  
 
  df_cc_est2016_20 <-
    df_cc_est2016_alldata_37 %>%
    dplyr::filter(
      YEAR == 6 &
      AGEGRP >= 5
    ) %>%  
    dplyr::select(
      -SUMLEV,
      -STATE,
      -STNAME,
      -CTYNAME,
      -YEAR
    )

  # AGEGRP == 4  

  df_cc_est2016_ <-
    df_cc_est2016_alldata_37 %>%
    dplyr::filter(
      YEAR == 6 &
      AGEGRP == 4
    ) %>%  
    dplyr::select(
      -SUMLEV,
      -STATE,
      -STNAME,
      -CTYNAME,
      -YEAR
    )

  # This computes 2/5 of ages 15-19 population  
 
  df_cc_est2016_tmp <- df_cc_est2016_ %>%
    dplyr::select(
      -COUNTY,
      -AGEGRP
    ) 
  df_cc_est2016_tmp[] <- round(0.4*df_cc_est2016_tmp[])

# Paste back together into a single data frame  
 
  df_cc_est2016_18 <-
    cbind(
      df_cc_est2016_ %>%
      dplyr::select(
        COUNTY,
        AGEGRP
      ),
      df_cc_est2016_tmp
    )
    
  # Combine to get all ages in one data frame  

  df_cc_est2016_voteage <-
    rbind(
      df_cc_est2016_20,
      df_cc_est2016_18
    ) %>%
    dplyr::arrange(
      COUNTY,
      AGEGRP
    )
    
  # Roll up to get total per county and add FIPS3  

  df_cc_est2016_voteagetot <-
    df_cc_est2016_voteage %>%
    dplyr::select(
      -AGEGRP
    ) %>%
    dplyr::group_by(
      COUNTY
    ) %>%
    dplyr::summarize_all(
      funs(sum)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      pop="voteage"
    )
  #
  # dataframe with all ages
  #
  df_cc_est2016_alltot <-
  df_cc_est2016_alldata_37  %>%
    dplyr::filter(
      YEAR==6 &
      AGEGRP==0
    ) %>%
    dplyr::select(
      -SUMLEV,
      -STATE,
      -STNAME,
      -CTYNAME,
      -YEAR,
      -AGEGRP
    ) %>%
    dplyr::mutate(
      pop="all"
    )
  #
  # combine
  #
  df_ <-
    rbind(
      df_cc_est2016_alltot,
      df_cc_est2016_voteagetot
    ) %>%
    dplyr::inner_join(
      df_FIPS %>% 
        dplyr::select(FIPS,FIPS3),
      by=c("COUNTY"="FIPS")
    )
  #
  return(df_)
}
```  

```{r sinew}
  df_cc_est2016_tot <- fn_df_cc_est2016_tot()
  str(df_cc_est2016_tot)
```  

Write out df_cc_est2016_tot as compressed file.  

```{r stove}
readr::write_csv(
  df_cc_est2016_tot,
  paste0(output_dir,"df_cc_est2016_tot.csv.gz")
)
```  

### County rollups  
WA White alone  
BA Black or African American alone  
IA American Indian and Alaska Native alone  
AA Asian alone  
NA Native Hawaiian and Other Pacific Islander alone  
TOM  Two or More Races  

Hispanic is double-counted but included as an additional column:  
H Hispanic  

These are double-counted and are not included:  
WAC  White alone or in combination  
BAC  Black or African American alone  
IAC  American Indian and Alaska Native alone or in combination  

The following function (data base view) works with df_cc_est2016_tot  
It can be used to generate a dataframe (as per SAS) or as an on-demand view.  

```{r angelo}
fn_df_cc_est2016_tot_subset <- function() {
  db_subset <-
  df_cc_est2016_tot %>%
  dplyr::select( 
    FIPS3,
    TOT_POP,
    TOT_MALE,
    TOT_FEMALE,
    WA_MALE,
    WA_FEMALE,
    BA_MALE,
    BA_FEMALE,
    IA_MALE,
    IA_FEMALE,
    AA_MALE,
    AA_FEMALE,
    NA_MALE,
    NA_FEMALE,
    TOM_MALE,
    TOM_FEMALE,
    WAC_MALE,
    WAC_FEMALE,
    BAC_MALE,
    BAC_FEMALE,
    IAC_MALE,
    IAC_FEMALE,
    AAC_MALE,
    AAC_FEMALE,
    H_MALE,
    H_FEMALE,
    pop
  ) %>%
  dplyr::mutate(
    nWA=WA_MALE+WA_FEMALE,
    nBA=BA_MALE+BA_FEMALE,
    nIA=IA_MALE+IA_FEMALE,
    nAA=AA_MALE+AA_FEMALE,
    nNA=NA_MALE+NA_FEMALE,
    nTOM=TOM_MALE+TOM_FEMALE,
    nH=H_MALE+H_FEMALE
  ) %>%
  dplyr::select(
    -TOT_MALE,
    -TOT_FEMALE,
    -WA_MALE,
    -WA_FEMALE,
    -BA_MALE,
    -BA_FEMALE,
    -IA_MALE,
    -IA_FEMALE,
    -AA_MALE,
    -AA_FEMALE,
    -NA_MALE,
    -NA_FEMALE,
    -TOM_MALE,
    -TOM_FEMALE,
    -WAC_MALE,
    -WAC_FEMALE,
    -BAC_MALE,
    -BAC_FEMALE,
    -IAC_MALE,
    -IAC_FEMALE,
    -AAC_MALE,
    -AAC_FEMALE,
    -H_MALE,
    -H_FEMALE
  )
  return(db_subset)
}
```  

```{r ash}  
df_cc_est2016_tot_subset <- fn_df_cc_est2016_tot_subset()
str(df_cc_est2016_tot_subset)
```  

Race by county:  

```{r gas}
df_cc_est2016_county_race_dcast <-
fn_df_cc_est2016_tot_subset() %>%
dplyr::select(TOT_POP,nWA,nBA,nIA,nAA,nTOM,pop,FIPS3) %>%
reshape2::melt(id.vars=c("pop","FIPS3")) %>%
reshape2::dcast(pop+FIPS3~variable,value.var = "value",sum)
```  

Convenience function to produce dataframe of male proportion:  

```{r spore}
fn_df_cc_est2016_mfprop <- function(){
# produces dataframe of
# proportion male based on df_cc_est2016_tot
#
  df_mf <- 
    df_cc_est2016_tot %>% 
    dplyr::select(
      TOT_POP:AAC_FEMALE,
      FIPS3,
      pop
    )
  #
  # get proportion male
  #
  df_mf %<>%
    dplyr::mutate(
      pmTOT=round(TOT_MALE/(TOT_MALE+TOT_FEMALE),digits=3),
      nTOT=TOT_MALE+TOT_FEMALE,
      pmWA=round(WA_MALE/(WA_MALE+WA_FEMALE),digits=3),
      nWA=WA_MALE+WA_FEMALE,    
      pmBA=round(BA_MALE/(BA_MALE+BA_FEMALE),digits=3),
      nBA=BA_MALE+BA_FEMALE,  
      pmIA=round(IA_MALE/(IA_MALE+IA_FEMALE),digits=3),
      nIA=IA_MALE+IA_FEMALE,  
      pmAA=round(AA_MALE/(AA_MALE+AA_FEMALE),digits=3),
      nAA=AA_MALE+AA_FEMALE,  
      pmTOM=round(TOM_MALE/(TOM_MALE+TOM_FEMALE),digits=3),
      nTOM=TOM_MALE+TOM_FEMALE
    ) %>%
    dplyr::select(
      pmTOT,
      pmWA,
      pmBA,
      pmIA,
      pmAA,
      pmTOM,
      nTOT,
      nWA,
      nBA,
      nIA,
      nAA,
      nTOM,
      FIPS3,
      pop
    ) 

  return(df_mf)
}
```  

```{r tong}
str(fn_df_cc_est2016_mfprop())
```  

## C. Precinct results file  
This has a lot of detail but says nothing about who voted.  

data source:  
https://dl.ncsbe.gov/index.html?prefix=ENRS/2016_11_08/  

```{r comic}
df_results_pct_20161108 <- read.csv(
  paste0(data_dir,"results_pct_20161108.txt"),
  sep="\t",
  header=TRUE,
  stringsAsFactors = FALSE
  )
```  

```{r pant}
str(df_results_pct_20161108)
```   

Trim all character columns  
Replace all blank or missing numerics with NA  
Mainly for Choice.Party  

```{r show}
df_results_pct_20161108_clean <- 
  df_results_pct_20161108 %>%
  mutate_if(
    is.character, funs(trimws(.)) 
  ) %>%
  base::replace(., .=="", NA)
```  

```{r tar}
str(df_results_pct_20161108_clean)
```  

Introduce some normalization  
Get distinct values for a few columns  
First, check that Contest.Name is independent of Contest.Type  

```{r vein}
fn_results2016_indep01 <- function() {
  df1 <- df_results_pct_20161108_clean %>%
    dplyr::distinct(
      Contest.Type,
      Contest.Name
    ) %>%
    dplyr::select(
      -Contest.Type
    ) %>%
    dplyr::group_by(
      Contest.Name
    ) %>%
    dplyr::filter(
      n()>1
    ) %>%
    dplyr::ungroup()
  return(df1)
}
#
# Stop Rmd execution if problem found
# Not sure if the View does anything
#
tmp_ <- nrow(fn_results2016_indep01())
if (tmp_!=0) {
  print(
    paste(
      "fn_results2016_indep01() should be empty but row count is",tmp_
    )
  )
  knitr::knit_exit()
}  
rm(tmp_)
```  
  
Now normalize  
"dfcl_" denotes a data frame that is a code list  

Contest.Name and Contest.Type are unambiguously paired.  

```{r wail}
dfcl_results_precinct_Contest.Type.Name <-
  dplyr::distinct(
    df_results_pct_20161108,
    Contest.Type,
    Contest.Name
  ) %>%
  dplyr::mutate(
    Contest.Type.Name.Code=as.numeric(rownames(.))
  )
#
dfcl_results_precinct_Choice <-
  dplyr::distinct(
    df_results_pct_20161108_clean,
    Choice
  ) %>%
  dplyr::mutate(
    Choice.Code=as.numeric(rownames(.))
  )
```  

```{r achoo}
str(dfcl_results_precinct_Contest.Type.Name)
str(dfcl_results_precinct_Choice)
```  

Drop meaningless trailing column X  
Replace County with FIPS3 code (not FIPS)  

```{r kulak}
df_results_pct_20161108_clean %<>%
  dplyr::inner_join(
    df_FIPS %>% 
      dplyr::select(
        County,
        FIPS3
      ) %>%
      dplyr::mutate(
        County=toupper(County)
      ),
    by="County") %>%
  dplyr::select(
    -County,
    -X
  )
```  

```{r moral}
str(df_results_pct_20161108_clean)
```  

Replace Contest.Type/Contest.Name with Contest.Type.Name.Code  

```{r fairy}
df_results_pct_20161108_clean %<>%
  dplyr::inner_join(
    dfcl_results_precinct_Contest.Type.Name,
    by=c("Contest.Type","Contest.Name")
  ) %>%
  dplyr::select(
    -Contest.Type,
    -Contest.Name
  )
```  

```{r mans}
str(df_results_pct_20161108_clean)
```  

Replace Choice with Choice.Code  

```{r price}
df_results_pct_20161108_clean %<>%
  dplyr::inner_join(
    dfcl_results_precinct_Choice,
    by="Choice") %>%
  dplyr::select(
    -Choice
  )
```  

```{r macho}
str(df_results_pct_20161108_clean)
```  

Simplify.

```{r dew}
df_results_pct_20161108_subset <-
  df_results_pct_20161108_clean %>%
  dplyr::select(
    FIPS3,
    Precinct,
    Choice.Party,
    Vote.For,
    Total.Votes,
    Contest.Type.Name.Code,
    Choice.Code
  )
```  

```{r lure}
str(df_results_pct_20161108_subset)
```  

## D. NCACC Elections spreadsheet  
Source is https://www.ncacc.org/409/County-Elections  
File is "2016 Election Demographics for Website_1_extract.csv"  
Skip row 1, which holds text for columns.  

```{r manic}
df_NCACC_2016_demog <- read.csv(  
  paste0(data_dir,"2016 Election Demographics for Website_1_extract.csv"),
  sep=",",
  skip=1,
  header=TRUE,
  stringsAsFactors=FALSE
)
```  

```{r dowel}
str(df_NCACC_2016_demog)
```  

Add FIPS3 and remove County  
Replace NA with zero  

```{r say}
df_NCACC_2016_demog %<>%
  dplyr::inner_join(
    df_FIPS,
    by="County"
  ) %>%
  dplyr::select(
    FIPS3,
    ncomm,
    new,
    nrep,
    ndem,
    nother,
    nfemale,
    nwhite,
    nafam,
    namin
  ) %>%
  dplyr::mutate_if(
    is.numeric, funs(ifelse(is.na(.), 0, .))
  )
```  

```{r oxnard}
str(df_NCACC_2016_demog)
```   

Write out df_NCACC_2016_demog as a compressed file.  

```{r tyler}
readr::write_csv(
  df_NCACC_2016_demog,
  paste0(output_dir,"df_NCACC_2016_demog.csv.gz")
)
``` 

## F. Finish  

```{r fly}
ls()
```  

```{r dose}
save.image("AWR_Census_NCACC.RData")
```  

### END
