---
title: "AreWeRepresented neo4j Simulation"
output:
  html_document: default
author: "D. Hopp"
date: "January 8, 2019"
---

```{r dyad}
load("prep.RData")
ls()
```

12/18/18 Change to replace=TRUE  

## Sim functions  

Run simulation by county based on df_totalbyrace  
Use only the three race categories reported in NCACC data plus Other:  
AmInd,Black,White,Other  

#
# sim functions  
#

```{r kapok}
fn_sim_byrace <- function(df_aug_,m_sim_) {
#
# run the sim n_sim times
# uses df_ACS_years_aug or df_totalbyrace_aug
# and m_ACS_years_sim or m_totalbyrace_sim
#
  n_row <- 0
  for (f_ in df_aug_$FIPS3) {
    for (n_ in 1:n_sim) {
      #
      # select a sample of size ncomm from between 1  
      #   and the total population of county
      #
      v_sim <- 
        sample.int(
          as.numeric(
            df_aug_[df_aug_$FIPS3==f_,"Total"]
          ),
          df_NCACC_demog[df_NCACC_demog$FIPS3==f_,"ncomm"],
          replace=TRUE
        )
      #
      # decide in which race category each sample falls
      #
      n_row <- n_row+1
      m_sim_[n_row,1] <- as.numeric(f_)
      for (s_ in 1:length(v_sim)) {
        if (v_sim[s_] <= df_aug_[df_aug_$FIPS3==f_,"AmIndEnd"]) {
          c <- 2
        } else if (v_sim[s_] <= df_aug_[df_aug_$FIPS3==f_,"BlackEnd"]) {
          c <- 3
        } else if (v_sim[s_] <= df_aug_[df_aug_$FIPS3==f_,"WhiteEnd"]){
          c <- 4
        } else {
          c <- 5
        }
        m_sim_[n_row,c] <- m_sim_[n_row,c] + 1
      }
    }
  }
  return(m_sim_)
}
```

```{r cozy}
fn_sim_counts <- function(m_sim_){
  #
  # Change m_ACS_years_sim to a data frame
  # produce df_ACS_years_sim_counts
  #
  df_sim <-
    as.data.frame.matrix(
      m_sim_
    )
  #
  #str(df_sim)
  #
  df_sim %<>%
    dplyr::inner_join(
      df_FIPS %>%
        dplyr::select(
          County,
          FIPS,
          FIPS3
        ),
      by=c("V1"="FIPS")
    ) %>%
    dplyr::rename(
      AmInd=V2,
      Black=V3,
      White=V4,
      Other=V5
    ) %>%
  dplyr::select(
      -V1
    ) %>%
    dplyr::group_by(
      AmInd,
      Black,
      White,
      Other,
      FIPS3
    )
  #
  #str(df_sim)
  #
  df_sim_counts <- 
    df_sim %>%
    dplyr::summarize(
      N=n()
    ) %>%
    dplyr::mutate(
      Npct=N*100/n_sim,
      Nlogit=ifelse(
        Npct==0,
        0,
        round(log(Npct/(100-Npct)),digits=4)
      )
    ) %>%
    dplyr::ungroup() %>%
    dplyr::arrange(
      FIPS3,
      -N
    )
  #
  return(df_sim_counts)
}
```

See if any of the actual NCACC race compositions are in the results of the sim.  
Use df_totalbyrace_sim_counts and df_NCACC_demog.  
Add on election method but in only three categories: at large, district, or mixed.  


```{r lumpy}
fn_sim_counts_hits <- function(df_sim_counts_){
  df_sim_counts_hits <-
  dplyr::left_join(
    df_NCACC_demog %>%
      dplyr::filter(
        Year==2016
      ) %>%
      dplyr::select(
        FIPS3,
        ncomm,
        nwhite,
        nafam,
        namin
      ),
    df_sim_counts_ %>%
      dplyr::filter(
        Other==0
      ),
    by=c(
      "FIPS3",
      "nwhite"="White",
      "nafam"="Black",
      "namin"="AmInd"
    )
  ) %>%
  dplyr::inner_join(
    df_FIPS %>%
      dplyr::select(
        FIPS3,
        County
      ),
    by="FIPS3"
  ) %>%
  dplyr::inner_join(
    df_Tier %>%
      dplyr::filter(
        Year==2017
      ) %>%
      dplyr::select(
        FIPS3,
        Tier
      ),
    by="FIPS3"
  ) %>%
  dplyr::inner_join(
    df_cc_method %>%
      dplyr::select(
        FIPS3,
        Method
      ),
    by="FIPS3"
  ) %>%
  dplyr::mutate(
    N=ifelse(is.na(N),0,N),
    Npct=ifelse(is.na(Npct),0,Npct),
    Nlogit=ifelse(
      Npct==0,
      0,
      round(log(Npct/(100-Npct)),digits=4)
    ),
    Method=ifelse(Method %in% c("al","d"),Method,"m")
  ) %>%
  dplyr::arrange(
    -N,
    County
  )
  #
  return(df_sim_counts_hits)
}
```  

A function to allow easy use of list return from a function.  
Source https://stackoverflow.com/questions/1826519/how-to-assign-from-a-function-which-returns-more-than-one-value  
Use:  
c(a,b) := someFunctionThatReturnsATwoItemList()  
a is the first item, b is the second.  

data.table has a := operator. May conflict???  

```{r flung}
':=' <- function(lhs, rhs) {
  frame <- parent.frame()
  lhs <- as.list(substitute(lhs))
  if (length(lhs) > 1)
    lhs <- lhs[-1]
  if (length(lhs) == 1) {
    do.call(`=`, list(lhs[[1]], rhs), envir=frame)
    return(invisible(NULL)) 
  }
  if (is.function(rhs) || is(rhs, 'formula'))
    rhs <- list(rhs)
  if (length(lhs) > length(rhs))
    rhs <- c(rhs, rep(list(NULL), length(lhs) - length(rhs)))
  for (i in 1:length(lhs))
    do.call(`=`, list(lhs[[i]], rhs[[i]]), envir=frame)
  return(invisible(NULL)) 
}
```  

This function adds results of a sim to the dataframes df_sim_results_counts, df_sim_results_hits and df_sim_results_index.  
df_sim_results_counts:  
"FIPS3" "AmInd" "Black" "White" "Other" "N" "Npct" "Nlogit" "index_uid"  

df_sim_results_hits (Method is election method):  
"FIPS3" "ncomm" "nwhite" "nafam" "namin" "Other" "N" "Npct" "Nlogit" "County" "Tier" "Method" "index_uid"  

df_sim_results_index
"index_uid" "Nsims" "Basis" "Election" "Dated" "Comment"  

Use a dataframe to keep track of index_uid. A constant might get lost. This has only one row.  
df_sim_results_next_index
"next_index_uid"  

```{r zinc}
fn_sim_result.add <- function(df_counts_,df_hits_,Nsims_,Basis_,Election_,Dated_,Comment_){
#
# Three dataframes are involved, one to hold counts, one to hold hits, one to act as an index.
# Check to see if dataframes exist. If none exist start anew. If some exist declare error.
#
# df_counts_ and df_hits_ are the results of a sim.
#
  check <- c(
    exists("df_sim_results_counts"),
    exists("df_sim_results_hits"),
    exists("df_sim_results_index"),
    exists("df_sim_results_next_index")
  )
  if (all(check)){
    #
    # add to existing data frames
    #
    new_index_uid <- df_sim_results_next_index[1,"next_index"]
    df_sim_results_counts <- 
      rbind(
        df_sim_results_counts,
        df_counts_ %>%
          dplyr::mutate(
            index_uid=new_index_uid
          )
      )
    #
    df_sim_results_hits <- 
      rbind(
        df_sim_results_hits,
        df_hits_ %>%
          dplyr::mutate(
            index_uid=new_index_uid
          )
      )
    #
    df_sim_results_index <-
      rbind(
        df_sim_results_index,
        c(
          as.numeric(Nsims_),
          as.character(Basis_),
          as.character(Election_),
          as.character(Dated_),
          as.character(Comment_),
          as.numeric(new_index_uid)
        )
      )
    df_sim_results_index$index_uid <- as.numeric(df_sim_results_index$index_uid)
    df_sim_results_index$Nsims <- as.numeric(df_sim_results_index$Nsims)
    #
    # update next_uid
    #
    df_sim_results_next_index[1,"next_index"] <- new_index_uid + 1
    #
  } else if (all(!check)) {
    #
    # create data frames
    # index_uid starts at 1
    #
    df_counts_ %<>% 
      dplyr::mutate(
        index_uid=1
      )
    df_sim_results_counts <- df_counts_
    #
    df_hits_ %<>% 
      dplyr::mutate(
        index_uid=1
      )
    df_sim_results_hits <- df_hits_
    #
    df_sim_results_index <-
      data.frame(
        Nsims_,
        Basis_,
        Election_,
        Dated_,
        Comment_
      ) %>%
      dplyr::rename(
        Nsims=Nsims_,
        Basis=Basis_,
        Election=Election_,
        Dated=Dated_,
        Comment=Comment_
      ) %>%
      dplyr::mutate(
        index_uid=1
      )
    #
    df_sim_results_index$index_uid <- as.numeric(df_sim_results_index$index_uid)
    df_sim_results_index$Nsims <- as.numeric(df_sim_results_index$Nsims)
    df_sim_results_index$Basis <- as.character(df_sim_results_index$Basis)
    df_sim_results_index$Election <- as.character(df_sim_results_index$Election)
    df_sim_results_index$Dated <- as.character(df_sim_results_index$Dated)
    df_sim_results_index$Comment <- as.character(df_sim_results_index$Comment)
    #
    df_sim_results_next_index <-
      data.frame(
        2
      )
    names(df_sim_results_next_index) <- "next_index"
    #
  } else {
    #
    # error that must be attended to
    # not all three dataframes are there
    #
    print(
      "ERROR: Only some of df_sim_results_counts, df_sim_counts_hits, df_sim_results_index and df_sim_results_next_index exist"
    )
    knitr::knit_exit()
  }
  #
  # finally
  #
  return(
    list(
      df_sim_results_counts,
      df_sim_results_hits,
      df_sim_results_index,
      df_sim_results_next_index
    )
  )
}
```  

```{r holly}
fn_sim_result.delete <- function(v_index_uid_){
#
# v_index_uid_ is a vector of positive integers
#
# delete records with any of v_index_uid in the dataframes 
# df_sim_results_counts, df_sim_results_hits and df_sim_results_index
#
# leaves df_sim_results_next_index unaltered
#
  check <- c(exists("df_sim_results_counts"),exists("df_sim_results_hits"),exists("df_sim_results_index"))
  if (all(check)){
    #
    # all data frames exist
    #
    df_sim_results_index %<>% 
      dplyr::filter(
        !(index_uid %in% v_index_uid_)
      )
    df_sim_results_counts %<>% 
      dplyr::filter(
        !(index_uid %in% v_index_uid_)
      )
    df_sim_results_index %<>% 
      dplyr::filter(
        !(index_uid %in% v_index_uid_)
      )
    #
    return(
      list(
        df_sim_results_counts,
        df_sim_results_hits,
        df_sim_results_index
      )
    )
  } else if (all(!check)) {
    #
    # none of the data frames are there
    #
    return(NULL)
  } else {
    #
    # some are there
    #
    print("Not all df_sim_results_counts, df_sim_results_hits and df_sim_results_index exist.")
    knitr::knit_exit()
  }
}
```

## Carry out sim  

```{r}
Sys.time()
```

Results will be placed in a matrix. We need a structure that holds for 
each county the sim results.  

Create a matrix with all 0 cells to hold the sim results  

Use fn_sim_result.add <- function(df_counts_,df_hits_,Nsims_,Basis_,Election_,Dated_,Comment_) to store results.  

Added August 8, 2018. Move away from use of df_totalbyrace. Simpler to use augmented df_ACS_years. Note however that population estimates can be substantially different.  

```{r shut}
df_ACS_years_aug <-
  df_ACS_years %>%
    dplyr::mutate(
      AmIndPct=round(AmInd*100/Total,digits=1),
      BlackPct=round(Black*100/Total,digits=1),
      WhitePct=round(White*100/Total,digits=1),      
      OtherPct=round(Other*100/Total,digits=1)
    )
#
str(df_ACS_years_aug)
#
# initialize save_core_sim_objects and save_this_sim_objects
#
save_core_sim_objects <- c("df_ACS_years","df_ACS_years_aug","df_totalbyrace","df_NCACC_demog")
save_this_sim_objects <- ""
```  

```{r way}
# AmInd can be zero
df_ACS_years_aug %<>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    AmIndEnd=AmInd,
    BlackEnd=AmIndEnd+Black,
    WhiteEnd=BlackEnd+White    
  ) %>%
  dplyr::ungroup()
#
str(df_ACS_years_aug)
```  

df_turnout_election_wide_aug is for comparing against voting population  

```{r jules}
#
# df_turnout_election_wide_aug is for comparing against census population
#
df_turnout_election_wide_aug <-
df_turnout_election_wide %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    Total=Black+AmInd+White+Other,
    AmIndEnd=AmInd,
    BlackEnd=AmIndEnd+Black,
    WhiteEnd=BlackEnd+White    
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    FIPS3,
    Election,
    AmIndEnd,
    BlackEnd,
    WhiteEnd,
    Total
  )
#
str(df_turnout_election_wide_aug)
save_core_sim_objects <- c(save_core_sim_objects,"df_turnout_election_wide_aug")
```  

Do the sim. Note especially the input to fn_sim_byrace(). If desired, name the data objects to match the sim data source and number of runs. Silently uses m_totalbyrace_sim.  

```{r krebs}
#
# set seed by inspiration or using runif BUT NOT BOTH!
#
# not really random - biased towards afternoon and evening
# use just decimal part
#
seed_1 <- 
    as.numeric(Sys.time())-as.numeric(Sys.Date())*86400
if (abs(round((seed_1-round(seed_1))*10000))==0){
  seed_ <- seed1_ + 1
} else {
  seed_ <- abs(round((seed_1-round(seed_1))*10000))
}
#
# or
#
#seed_ <- 1234
#
set.seed(seed_)
#
# runs per county:
n_sim <- 2000
#
save_this_sim_objects <- c(save_this_sim_objects,"n_sim","seed_")
save_file_suffix <- ""
#
seed_
n_sim
```  

Sim for census data and 2016 general election  

```{r juror}
#
# DO NOT USE
#
do_census_161108 <- FALSE
#
if (do_census_161108) {
  #
  # first column will be numeric interpretation of FIPS3
  # then AmInd,Black,White,Other
  #
  m_totalbyrace_sim <-
    matrix(
      rep.int(0,100*n_sim*5),
      ,
      nrow=100*n_sim,
      ncol=5
    )
  #
  # does the sim
  #
  m_totalbyrace_sim <- 
    fn_sim_byrace(
      df_totalbyrace_aug,
      m_totalbyrace_sim
    )
  str(m_totalbyrace_sim)
  #
  # uses m_totalbyrace_sim
  #
  df_totalbyrace_sim_counts_census_161108_ <- 
    fn_sim_counts(m_totalbyrace_sim)
  str(df_totalbyrace_sim_counts_census_161108_)
  #
  df_totalbyrace_sim_counts_hits_census_161108_ <- 
    fn_sim_counts_hits(
      df_totalbyrace_sim_counts_census_161108_
    )
  str(df_totalbyrace_sim_counts_hits_census_161108_)
  #
  # save in an RData file and remove from objects
  #
  save(
    df_totalbyrace_sim_counts_hits_census_161108_,
    df_totalbyrace_sim_counts_census_161108_,
    file="totalbyrace_sim_counts_census_161108_.RData"
  )
  rm(
    df_totalbyrace_sim_counts_hits_census_161108_,
    df_totalbyrace_sim_counts_census_161108_
  )
  save(
    m_totalbyrace_sim,
    file="m_totalbyrace_sim_census_161108_.RData"
  )
  rm(m_totalbyrace_sim)
  #
  save_file_suffix <- paste0(save_file_suffix,"_total")
}
```  

Sim for ACS data and 2016 general election  
NOTE: For now must change by hand object names ending in number of sims  

```{r want}
do_ACS_161108 <- FALSE
#
if (do_ACS_161108) {
  #
  # first column will be numeric interpretation of FIPS3
  # then AmInd,Black,White,Other
  #
  m_ACS_years_sim <-
    matrix(
      rep.int(0,100*n_sim*5),
      ,
      nrow=100*n_sim,
      ncol=5
    )
  #
  # does the sim
  #
  m_ACS_years_sim <- 
    fn_sim_byrace(
      df_ACS_years_aug %>% 
        dplyr::filter(Year=="2016"),
      m_ACS_years_sim
    )
  str(m_ACS_years_sim)
  #
  # uses m_ACS_years_sim
  #
  df_ACS_years_sim_counts_census_161108_ <- 
    fn_sim_counts(m_ACS_years_sim)
  #
  df_ACS_years_sim_counts_census_161108_ %<>%
    dplyr::mutate(
      n_sim=n_sim,
      seed=seed_
    )
  #
  str(df_ACS_years_sim_counts_census_161108_)
  #  
  df_ACS_years_sim_counts_hits_census_161108_ <- 
    fn_sim_counts_hits(
      df_ACS_years_sim_counts_census_161108_
    )
  #
  str(df_ACS_years_sim_counts_hits_census_161108_)
  #
  # add explicit row number for later use
  #
  df_ACS_years_sim_counts_census_161108_$n_row <-
    as.numeric(
      row.names(
        df_ACS_years_sim_counts_census_161108_
      )
    )  
  #
  # change names and save as RData 
  #
  obj1 <- paste0("df_ACS_years_sim_counts_census_161108_",n_sim,"_rep_",seed_)
  obj2 <- paste0("df_ACS_years_sim_counts_hits_census_161108_",n_sim,"_rep_",seed_)
  #
  assign(
    obj1,
    df_ACS_years_sim_counts_census_161108_
  )
  assign(
    obj2,
    df_ACS_years_sim_counts_hits_census_161108_
  )  
  #  
  save_this_sim_objects <- c(save_this_sim_objects,
    obj1,
    obj2
  )  
  #
  save_file_suffix <- paste0(save_file_suffix,"_ACS")
  #
  rm(m_ACS_years_sim)
}
```  

```{r}
Sys.time()
```

Sim for voting data - NCSBE counts of persons who voted in 2016 general election  

```{r edith}
do_voting_161108 <- TRUE
#
if (do_voting_161108) {
  #
  # first column will be numeric interpretation of FIPS3
  # then AmInd,Black,White,Other
  #
  m_voting_sim <-
    matrix(
      rep.int(0,100*n_sim*5),
      ,
      nrow=100*n_sim,
      ncol=5
    )
  #
  # does the sim
  #
  m_voting_sim <- 
    fn_sim_byrace(
      df_turnout_election_wide_aug %>% 
        dplyr::filter(
          Election=="20161108"
        ),
      m_voting_sim
    )
  str(m_voting_sim)
  #
  # uses m_voting_sim
  #
  df_voting_sim_counts_ <- 
    fn_sim_counts(
      m_voting_sim
    )
  #
  df_voting_sim_counts_ %<>%
    dplyr::mutate(
      n_sim=n_sim,
      seed=seed_
    )
  #
  str(df_voting_sim_counts_)
  #
  df_voting_sim_counts_hits_ <- 
    fn_sim_counts_hits(
      df_voting_sim_counts_
    )
  #
  str(df_voting_sim_counts_hits_)
  #
  # add explicit row number for later use
  #
  df_voting_sim_counts_$n_row <-
    as.numeric(
      row.names(
        df_voting_sim_counts_
      )
    )
  #
  c(df_sim_results_counts,
    df_sim_results_hits,
    df_sim_results_index,
    df_sim_results_next_index
  ) := fn_sim_result.add(
    df_voting_sim_counts_,
    df_voting_sim_counts_hits_,
    n_sim,
    "turnout",
    "20161108",
    format(Sys.Date(), format="%Y%m%d"),
    ""
  )
  #
  rm(m_voting_sim)
  #
  # change names and save
  #
  obj1 <- paste0("df_voting_sim_counts_",n_sim,"_rep_",seed_)
  obj2 <- paste0("df_voting_sim_counts_hits_",n_sim,"_rep_",seed_)
  #
  assign(
    obj1,
    df_voting_sim_counts_
  )
  assign(
    obj2,
    df_voting_sim_counts_hits_
  )
  #
  save_this_sim_objects <- c(save_this_sim_objects,
    obj1,
    obj2
  )
  #
  save_file_suffix <- paste0(save_file_suffix,"_voting")
}
```  

```{r}
Sys.time()
```

```{r savoy}
ls()
#
# save the save-lists
#
save_core_sim_objects <- c(save_core_sim_objects,"save_core_sim_objects")
#
# display it
#
save_core_sim_objects
#
# do the save
#
save(
  list=save_core_sim_objects,
  file=paste0("core_sim_objects_",n_sim,save_file_suffix,"_rep_",seed_,".RData")
)
```

```{r savoy2}
#
save_this_sim_objects <- c(save_this_sim_objects,"save_this_sim_objects")
#
# display it
#
save_this_sim_objects
#
# do the save
#
save(
  list=save_this_sim_objects,
  file=paste0(output_dir,"this_sim_objects_",n_sim,save_file_suffix,"_rep_",seed_,".RData")
)
```

```{r eval=FALSE, echo=FALSE}
#
# write to csv
#
for (s in save_this_sim_objects){
  if (any(class(get(s))=="data.frame")){
    readr::write_delim(
      get(s),
      path=paste0(output_dir,s,".csv"),
      delim="\t"
    )
  }
}
```  
