---
title: "AreWeRepresented neo4j Jeremy Simulation"
output:
  html_document: default
author: "D. Hopp"
date: "December 22, 2018"
---

Revised 12/13/18  

```{r dyad}
load("prep.RData")
ls()
```


Read Jeremy set of "result" files, one per county  
File name includes county code of form "ccc.result"  
where ccc is the FIPS code  
Files are of form:  
{('AMI', 'AMI', 'OTR', 'WHT', 'WHT'): 1,  
 ('AMI', 'AMI', 'WHT', 'WHT', 'WHT'): 3,  
 ('AMI', 'BLK', 'WHT', 'WHT', 'WHT'): 6,  
 ('AMI', 'OTR', 'WHT', 'WHT', 'WHT'): 12,  
 ('AMI', 'WHT', 'WHT', 'WHT', 'WHT'): 177,  
 ('BLK', 'BLK', 'WHT', 'WHT', 'WHT'): 13,  
 ('BLK', 'OTR', 'WHT', 'WHT', 'WHT'): 40,  
 ('BLK', 'WHT', 'WHT', 'WHT', 'WHT'): 486,  
 ('OTR', 'OTR', 'WHT', 'WHT', 'WHT'): 32,  
 ('OTR', 'WHT', 'WHT', 'WHT', 'WHT'): 821,  
 ('WHT', 'WHT', 'WHT', 'WHT', 'WHT'): 8409}  

Handy-hint:  
https://stackoverflow.com/questions/12626637/reading-a-text-file-in-r-line-by-line  



```{r }
fn_read_jeremy_as_one_file <- function(){
  #
  n_sim <- 10000
  #
  # create a matrix to hold composite results from all Jeremy files
  # determine rows of matrix by scanning all the files
  # first set up helper dataframe
  # convert to dataframe
  #
  df_F <- dplyr::inner_join(
    df_FIPS,
    df_NCACC_years %>%
      dplyr::filter(
        Year=="2016"
      ) %>%
      dplyr::select(
        County,
        ncomm
      ),
    by="County"
  ) %>%
  dplyr::arrange(
    ncomm,
    FIPS
  )
  #
  file_dir <- "../AWR_data_only/ACS_2016_10k_sim/"
  #
  nrow_matrix <- 0
  for (FIPS_ in df_F[,"FIPS"]){
    com <- paste("wc -l ", file_dir, FIPS_, ".result", sep="")
    ntext  <- system(command=com, intern=TRUE)
    nrow_matrix <- nrow_matrix + as.integer(stringr::str_sub(ntext,1,stringr::str_locate(ntext," ")[1]-1))
  }
  #
  # create matrix
  # columns AmInd,Black,White,Other,ncomm,N,FIPS as
  # B,I,O,W,ncomm,N,FIPS
  #
  m_ <- matrix(
    rep(0,7),
    nrow=nrow_matrix,
    ncol=7
  )
  colnames(m_)=c("AmInd","Black","White","Other","ncomm","N","FIPS")
  #
  # iterate thru files
  #
  m_row <- 0
  for (fr_ in 1:nrow(df_F)){
    FIPS_ <- df_F[fr_,"FIPS"]
    nc_ <- df_F[fr_,"ncomm"]
    # specify the input file
    file_ <- paste0(
      file_dir,FIPS_,".result"
    )
    # create a connection
    conn <- file(
      description=file_, 
      open="r"
    )
    # slurp it
    linn <- readLines(conn)
    n_ <- length(linn)
    #
    # Loop over the file lines
    for(r_ in 1:n_) {
      tmp <- linn[r_]
      # do something on a line of data 
      nB <- stringr::str_count(tmp,"BLK")
      nI <- stringr::str_count(tmp,"AMI")
      nO <- stringr::str_count(tmp,"OTR")
      nW <- stringr::str_count(tmp,"WHT")
      N_ <- stringr::str_extract(tmp,'\\d+')
      #
      # construct vector of values and write line to m_
      #
      m_row <- m_row+1
      m_[m_row,1:7] <- c(
        nI,
        nB,
        nW,
        nO,
        nc_,
        N_,
        FIPS_
      )
    }    
    close.connection(conn)
  }
  #
  # convert to dataframe
  #
  df_ <- data.frame(
    m_,
    stringsAsFactors=FALSE
  ) %>%
  dplyr::mutate_if(
    is.character,
    as.numeric
  ) %>%
  dplyr::inner_join(
    df_FIPS %>%
      dplyr::select(
        FIPS,
        FIPS3
      ),
    by="FIPS"
  ) %>%
  dplyr::select(
    -FIPS
  ) %>%
  dplyr::arrange(
    FIPS3,
    -N
  )
  #
  # add Npct and Nlogit
  # and running sum of Npct
  #
  df_ %<>%
    dplyr::mutate(
      Npct=N*100/n_sim,
      Nlogit=ifelse(
        Npct==0,
        0,
        round(log(Npct/(100-Npct)),digits=4)
      )
    ) %>%
    dplyr::group_by(
      FIPS3
    ) %>%
    dplyr::ungroup()
  #
  # add columns used by other sims  
  # seed unknown
  #
  df_ %<>%
    dplyr::mutate(
      n_sim=10000,
      seed=0
    ) 
  #
  # add row number
  #
  df_$n_row <-
    as.numeric(
      row.names(
        df_
      )
    ) 
  #
  # set column order as per other sims
  #
  df_%<>%
    dplyr::select(
      AmInd,
      Black,
      White,
      Other,
      FIPS3,
      N,
      Npct,
      Nlogit,
      n_sim,
      seed,
      n_row
    )
  #
  return(df_)
}
```

Preferred name of Jeremy sim:  

```{r }
df_ACS_sim_jf_10k <- fn_read_jeremy_as_one_file()
```

Create hits dataframe with rows matching board composition.  
Use 2016 as the year.  

```{r }
fn_jeremy_hits <- function(use_dataframe=FALSE){
  #
  # value: hits dataframe derived from df_ACS_sim_jf_10k
  #
  # use_dataframe=FALSE
  #   use fn_read_jeremy_as_one_file()
  # TRUE use dataframe df_ACS_sim_jf_10k
  #
  if (use_dataframe){
    df_N <- df_ACS_sim_jf_10k
  } else {
    df_N <- fn_read_jeremy_as_one_file()
  }
  #
  # add FIPS and FIPS3 to df_NCACC_years
  #
  df_cc <- df_NCACC_years %>%
    dplyr::filter(
      Year=="2016"
    ) %>%
    dplyr::inner_join(
      df_FIPS %>%
        dplyr::select(
          County,
          FIPS,
          FIPS3
        ),
      by="County"
    )
  #
  # trivial Cartesian product since df_cc is 1:1 with counties
  #
  df_hits <-
    dplyr::left_join(
      df_cc,
      df_N,
      by="FIPS3"
    )
  #
  # select the matching line
  #
  df_hits %<>%
    dplyr::filter(
      nwhite==White &
      nafam==Black &
      namin==AmInd
    ) %>%
    dplyr::select(
      FIPS3,
      nwhite,
      nafam,
      namin,
      Other,
      N,
      Npct,
      County
    )
  #
  return(df_hits)
}
```

Preferred name of hits (matches to 2016 board composition)  

```{r }
df_ACS_sim_hits_jf_10k <- fn_jeremy_hits()
```

```{r }
save(
  list=c(
    "df_ACS_sim_jf_10k",
    "df_ACS_sim_hits_jf_10k"
  ),
  file="ACS_sim_objects_jf_10k.RData"
)
```
