---
title: "AreWeRepresented Sim Clustering"
output:
  html_document: default
author: "D. Hopp"
date: "January 7, 2019"
---

```{r }
load("core_sim_objects.RData")
#load("sim_objects_1000.RData")
ls()
```

```{r}
# https://stackoverflow.com/questions/1826519/how-to-assign-from-a-function-which-returns-more-than-one-value?noredirect=1&lq=1
# c(a, b) := functionReturningTwoValues()
#
':=' <- function(lhs, rhs) {
  frame <- parent.frame()
  lhs <- as.list(substitute(lhs))
  if (length(lhs) > 1)
    lhs <- lhs[-1]
  if (length(lhs) == 1) {
    do.call(`=`, list(lhs[[1]], rhs), envir=frame)
    return(invisible(NULL)) 
  }
  if (is.function(rhs) || is(rhs, 'formula'))
    rhs <- list(rhs)
  if (length(lhs) > length(rhs))
    rhs <- c(rhs, rep(list(NULL), length(lhs) - length(rhs)))
  for (i in 1:length(lhs))
    do.call(`=`, list(lhs[[i]], rhs[[i]]), envir=frame)
  return(invisible(NULL)) 
}
```
## Clustering Analysis of the Sim Results  

Use package Ckmeans.1d.dp  
Note that weights cannot be used.  

Can use vector of race-board member counts interpreted as a number  
but package wants sorted data so that will not work.  

Could assign the row number in a sorted sim results data frame and transform  
into n_sim rows (each row a single instance of the row number).  

How will the clustering algorithm act with so many repetitions?  

names(df_voting_sim_counts_1000)  
[1] "AmInd"  "Black"  "White"  "Other"  "FIPS3"  "N"      "Npct"   "Nlogit"  

```{r }
fn_do_Ckmeans.1d.dp <- function(sim,k_=4){
  #
  # sim is the results of a simulation run, e.g., df_voting_sim_counts_1000
  # k_ is the starting (max) number of clusters to look for
  #
  # if Ckmeans.1d.dp is satisfied with k_ it stops
  # if not it decreases k_ to a min value of 2 before giving up
  #
  # this returns a dataframe with one row cluster for each county
  # the row contains:
  #   the FIPS3 code as a numeric, 
  #   the center of a cluster,
  #   the withinss (sum of squares of deviations)
  #   the size is number of rows of the input matrix that belong to the cluster
  #
  fn_transform_sim <- function(){
    #
    # sim is the results of a simulation run, e.g., df_voting_sim_counts_1000
    #
    # returns a matrix that can be used for further analysis
    # the matrix has one row per instance of the sim result
    # in the returned matrix:
    # FIPS3 is the numeric county FIPS3
    # n is just a row number that refers to the INPUT sim
    # N is the count associated with that row number in the input sim
    #  it is repeated N times which means that the race counts in
    #  row n would be repeated that many times
    # 
#browser()
    df_transform_ <-
      sim %>%
      dplyr::select(
        FIPS3,
        N,
        n_row
      )
    #
    # use FIPS3_ as a group indicator
    # and n_row as an index
    #
    m <- matrix(
      nrow=sum(df_transform_$N),
      ncol=3
    )
    m_row <- 0
    for (i in 1:nrow(df_transform_)) {
      FIPS3_ <- as.numeric(df_transform_[i,"FIPS3"])
      n_ <- as.numeric(df_transform_[i,"n_row"])
      N_ <- as.numeric(df_transform_[i,"N"])
      for (j in 1:N_) {
        m_row <- m_row+1
        m[m_row,1] <- FIPS3_
        m[m_row,2] <- n_
        m[m_row,3] <- N_
      }
    }
    colnames(m) <- c("FIPS3_","n_row","N")
    #
    # use colnames() to retrieve names, not names()
    #
    return(m)
  }  
  # end function def
  # ############################
  #
  # begin computations
  #
  # mat is created from the function fn_transform_sim
  #
#browser()
  mat <- fn_transform_sim()
  #
  county_ <- unique(mat[,1])
  #
  # assumes same number of rows for all counties
  #
  chunk_ <- as.integer(nrow(mat)/length(county_))
  # check
  if (nrow(mat) != chunk_*length(county_)) {
    print(
      paste(
        "Error: chunk count seems to be wrong.",
        "\nnrow(mat) =", nrow(mat),
        " chunk =",chunk
      )
    )
    knitr::knit_exit()
  }
  #
  # df_results will hold all the Ckmeans.1d.dp objects
  #
  n_res <- 1
  n_mat <- 1
  #
  for (c_ in county_){
    trythis_ <- TRUE
    kthis_ <- k_
    while (trythis_){
      res <- Ckmeans.1d.dp(
        mat[as.integer(n_mat):as.integer(n_mat+chunk_-1),3],
        k=kthis_
      )
      #
      # see if successful and whether to try again
      #
      kthis_ <- kthis_-1
      trythis_ <- (length(res[[1]])==1 & kthis_>1)
    } # end while
    #
#browser()
    if (!exists("df_clusters",inherits=FALSE)) {
      df_clusters <- data.frame(
        FIPS3_=c_,
        res2=res[2],
        res3=res[3],
        res4=res[4],
        stringsAsFactors=FALSE
      )
    } else {
      tmp_ <- data.frame(
        FIPS3_=c_,
        res2=res[2],
        res3=res[3],
        res4=res[4],
        stringsAsFactors=FALSE
      )
      df_clusters <- rbind(
        df_clusters,
        tmp_
      )
      rm(tmp_)
    }
    #
    n_res <- n_res+1
    n_mat <- n_mat+chunk_
  }
  # 
  # sort to match order of "sim" and
  # assign an explicit cluster number
  # (the combination of FIPS3_ and n_clus)
  #
  df_clusters %<>%
    dplyr::mutate(
      centers=round(centers,digits=1),
      withinss=round(withinss,digits=1)
    ) %>%
    dplyr::arrange(
      FIPS3_,
      -centers
    ) %>%
    dplyr::group_by(
      FIPS3_
    ) %>% 
    dplyr::mutate(
      n_clus = row_number()
    ) %>%
    dplyr::ungroup()
  #
  # which rows of sim do the clusters include?
  # this combines n_row in "sim" and "size" in df_clusters,
  # connected by FIPS3_ and n_clus
  #
  # each row in df_cluster_row will be associated with
  # either 0 or 1 cluster
  #
  df_cluster_row <-
    sim %>%
    dplyr::select(
      FIPS3,
      N,
      n_row
    ) %>%
    dplyr::mutate(
      FIPS3_=as.integer(FIPS3),
      n_clus=0
    )
  #   
  # row of df_clusters
  r_clusters <- 1
  running_sum <- 0
  #
  # step through df_clusters keeping track of row as r_
  # df_clusters has one row per cluster while df_cluster_row
  # is finer grained
  #
  # task is to assign a cluster (from df_clusters) to each
  # row of df_cluster_row
  #
  for (r_ in 1:nrow(df_cluster_row)){
    #
    thisFips <- as.numeric(df_cluster_row[r_,"FIPS3_"])
    # pick "size" from df_clusters
    size_ <- as.numeric(df_clusters[r_clusters,"size"])
    # and N from df_cluster_row
    N_ <- as.numeric(df_cluster_row[r_,"N"])
    # compare to this, not to N_:
    running_sum <- running_sum + N_
    #
    # if exact match of "size" to N, then immediately assign cluster (n_clus)
    # and see what comes next
    # but if not an exact fit, then work through df_cluster_row until
    # running N sum equals "size"
    #
    if (size_ == running_sum) {
      # good fit
      df_cluster_row[r_,"n_clus"] <- as.numeric(df_clusters[r_clusters,"n_clus"])
      running_sum <- 0
      # move to next row of df_clusters
      r_clusters <- r_clusters + 1
      #
      if (r_clusters > nrow(df_clusters)+1){
        print(paste("ERROR: overrun of df_clusters for r_:",r_))
        knitr::knit_exit()
      }
    # size_ != running_sum
    } else {
      # stick it 
      df_cluster_row[r_,"n_clus"] <- as.numeric(df_clusters[r_clusters,"n_clus"])
      # DO NOT move to next row of df_clusters
    }
  }
  #
  # swap FIPS3 for FIPS3_
  # should have done this earlier
  #
  df_clusters %<>%
    dplyr::inner_join(
      df_FIPS %>%
        dplyr::select(
          FIPS,
          FIPS3
        ),
      by=c("FIPS3_"="FIPS")
    ) %>%
    dplyr::select(
      -FIPS3_
    )
  #
  df_cluster_row %<>%
    dplyr::select(
      -FIPS3_
    )
  #
  # done
  #  
  return(
    list(
      df_clusters,
      df_cluster_row
    )
  )
}
```

```{r }
c(
  df_clusters_voting_1000_5,
  df_cluster_row_voting_1000_5
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_1000,
    k_=5
  )
#
c(
  df_clusters_ACS_161108_1000_5,
  df_cluster_row_ACS_161108_1000_5
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_ACS_years_sim_counts_census_161108_1000,
    k_=5
  )
```

### New  

```{r }
#
# 1264
#
c(
  df_cluster_5_voting_2000_rep_1264,
  df_cluster_5_row_voting_2000_rep_1264
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_2000_rep_1264,
    k_=5
  )
#
df_cluster_5_voting_2000_rep_1264 %<>%
  dplyr::mutate(
    k="5",
    n_sim="2000",
    replace=TRUE,
    seed="1264"
  )
#
df_cluster_5_row_voting_2000_rep_1264 %<>%
  dplyr::mutate(
    k="5",
    n_sim="2000",
    replace=TRUE,
    seed="1264"
  )
#
c(
  df_cluster_3_voting_2000_rep_1264,
  df_cluster_3_row_voting_2000_rep_1264
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_2000_rep_1264,
    k_=3
  )
#
df_cluster_3_voting_2000_rep_1264 %<>%
  dplyr::mutate(
    k="3",
    n_sim="2000",
    replace=TRUE,
    seed="1264"
  )
#
df_cluster_3_row_voting_2000_rep_1264 %<>%
  dplyr::mutate(
    k="3",
    n_sim="2000",
    replace=TRUE,
    seed="1264"
  )
#
# 1677
#
c(
  df_cluster_5_voting_2000_rep_1677,
  df_cluster_5_row_voting_2000_rep_1677
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_2000_rep_1677,
    k_=5
  )
#
df_cluster_5_voting_2000_rep_1677 %<>%
  dplyr::mutate(
    k="5",
    n_sim="2000",
    replace=TRUE,
    seed="1677"
  )
#
df_cluster_5_row_voting_2000_rep_1677 %<>%
  dplyr::mutate(
    k="5",
    n_sim="2000",
    replace=TRUE,
    seed="1677"
  )
#
c(
  df_cluster_3_voting_2000_rep_1677,
  df_cluster_3_row_voting_2000_rep_1677
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_2000_rep_1677,
    k_=3
  )
#
df_cluster_3_voting_2000_rep_1677 %<>%
  dplyr::mutate(
    k="3",
    n_sim="2000",
    replace=TRUE,
    seed="1677"
  )
#
df_cluster_3_row_voting_2000_rep_1677 %<>%
  dplyr::mutate(
    k="3",
    n_sim="2000",
    replace=TRUE,
    seed="1677"
  )
#
# 2935
# 
c(
  df_cluster_5_voting_2000_rep_2935,
  df_cluster_5_row_voting_2000_rep_2935
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_2000_rep_2935,
    k_=5
  )
#
df_cluster_5_voting_2000_rep_2935 %<>%
  dplyr::mutate(
    k="5",
    n_sim="2000",
    replace=TRUE,
    seed="2935"
  )
#
df_cluster_5_row_voting_2000_rep_2935 %<>%
  dplyr::mutate(
    k="5",
    n_sim="2000",
    replace=TRUE,
    seed="2935"
  )
#
c(
  df_cluster_3_voting_2000_rep_2935,
  df_cluster_3_row_voting_2000_rep_2935
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_2000_rep_2935,
    k_=3
  )
#
df_cluster_3_voting_2000_rep_2935 %<>%
  dplyr::mutate(
    k="3",
    n_sim="2000",
    replace=TRUE,
    seed="2935"
  )
#
df_cluster_3_row_voting_2000_rep_2935 %<>%
  dplyr::mutate(
    k="3",
    n_sim="2000",
    replace=TRUE,
    seed="2935"
  )
#
# 3226
#
c(
  df_cluster_5_voting_2000_rep_3226,
  df_cluster_5_row_voting_2000_rep_3226
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_2000_rep_3226,
    k_=5
  )
#
df_cluster_5_voting_2000_rep_3226 %<>%
  dplyr::mutate(
    k="5",
    n_sim="2000",
    replace=TRUE,
    seed="3226"
  )
#
df_cluster_5_row_voting_2000_rep_3226 %<>%
  dplyr::mutate(
    k="5",
    n_sim="2000",
    replace=TRUE,
    seed="3226"
  )
#
c(
  df_cluster_3_voting_2000_rep_3226,
  df_cluster_3_row_voting_2000_rep_3226
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_2000_rep_3226,
    k_=3
  )
#
df_cluster_3_voting_2000_rep_3226 %<>%
  dplyr::mutate(
    k="3",
    n_sim="2000",
    replace=TRUE,
    seed="3226"
  )
#
df_cluster_3_row_voting_2000_rep_3226 %<>%
  dplyr::mutate(
    k="3",
    n_sim="2000",
    replace=TRUE,
    seed="3226"
  )
#
```

Jeremy ACS sim.  

```{r }
c(
  df_cluster_5_jf_10k_rep,
  df_cluster_5_row_jf_10k_rep
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_ACS_sim_jf_10k,
    k_=5
  )
#
df_cluster_5_jf_10k_rep %<>%
  dplyr::mutate(
    k="5",
    n_sim="10000",
    replace=TRUE,
    seed=""
  ) 
#
df_cluster_5_row_jf_10k_rep %<>%
  dplyr::mutate(
    k="5",
    n_sim="10000",
    replace=TRUE,
    seed=""
  )
#
c(
  df_cluster_3_jf_10k_rep,
  df_cluster_3_row_jf_10k_rep
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_ACS_sim_jf_10k,
    k_=3
  )
#
df_cluster_3_jf_10k_rep %<>%
  dplyr::mutate(
    k="3",
    n_sim="10000",
    replace=TRUE,
    seed=""
  )
#
df_cluster_3_row_jf_10k_rep %<>%
  dplyr::mutate(
    k="3",
    n_sim="10000",
    replace=TRUE,
    seed=""
  )
```

5000 run voting sim  

```{r }
c(
  df_cluster_5_voting_5000_rep_1655,
  df_cluster_5_row_voting_5000_rep_1655
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_5000_rep_1655,
    k_=5
  )
#
df_cluster_5_voting_5000_rep_1655 %<>%
  dplyr::mutate(
    k="5",
    n_sim="5000",
    replace=TRUE,
    seed="1655"
  ) 
#
df_cluster_5_row_voting_5000_rep_1655 %<>%
  dplyr::mutate(
    k="5",
    n_sim="5000",
    replace=TRUE,
    seed="1655"
  )
#
c(
  df_cluster_3_voting_5000_rep_1655,
  df_cluster_3_row_voting_5000_rep_1655
  ) :=
  fn_do_Ckmeans.1d.dp(
    df_voting_sim_counts_5000_rep_1655,
    k_=3
  )
#
df_cluster_3_voting_5000_rep_1655 %<>%
  dplyr::mutate(
    k="3",
    n_sim="5000",
    replace=TRUE,
    seed="1655"
  )
#
df_cluster_3_row_voting_5000_rep_1655 %<>%
  dplyr::mutate(
    k="3",
    n_sim="5000",
    replace=TRUE,
    seed="1655"
  )
```

### IMPORTANT - DO THIS FOR VOTING SIMS  

Meld into supersets and get rid of lots of objects.  
Add time stamp.  

```{r }
dated <- Sys.time()
df_cluster_voting_all <-
  rbind(
    df_cluster_5_voting_2000_rep_1264,
    df_cluster_3_voting_2000_rep_1264,
    df_cluster_5_voting_2000_rep_1677,
    df_cluster_3_voting_2000_rep_1677,
    df_cluster_5_voting_2000_rep_2935,
    df_cluster_3_voting_2000_rep_2935,
    df_cluster_5_voting_2000_rep_3226,
    df_cluster_3_voting_2000_rep_3226
  ) %>%
  dplyr::mutate(
    Dated=dated
  )
#
df_cluster_row_voting_all <-
  rbind(
    df_cluster_5_row_voting_2000_rep_1264,
    df_cluster_3_row_voting_2000_rep_1264,
    df_cluster_5_row_voting_2000_rep_1677,
    df_cluster_3_row_voting_2000_rep_1677,
    df_cluster_5_row_voting_2000_rep_2935,
    df_cluster_3_row_voting_2000_rep_2935,
    df_cluster_5_row_voting_2000_rep_3226,
    df_cluster_3_row_voting_2000_rep_3226
  ) %>%
  dplyr::mutate(
    Dated=dated
  )
```

Archive   

```{r }
save(
    df_cluster_5_voting_2000_rep_1264,
    df_cluster_3_voting_2000_rep_1264,
    df_cluster_5_voting_2000_rep_1677,
    df_cluster_3_voting_2000_rep_1677,
    df_cluster_5_voting_2000_rep_2935,
    df_cluster_3_voting_2000_rep_2935,
    df_cluster_5_voting_2000_rep_3226,
    df_cluster_3_voting_2000_rep_3226,
    df_cluster_5_row_voting_2000_rep_1264,
    df_cluster_3_row_voting_2000_rep_1264,
    df_cluster_5_row_voting_2000_rep_1677,
    df_cluster_3_row_voting_2000_rep_1677,
    df_cluster_5_row_voting_2000_rep_2935,
    df_cluster_3_row_voting_2000_rep_2935,
    df_cluster_5_row_voting_2000_rep_3226,
    df_cluster_3_row_voting_2000_rep_3226,
    file=paste0(output_dir,"archive_cluster_row_voting_2000_all_A.RData")
  )
```  

Add the 5000 run sim  

```{r }
dated <- Sys.time()
df_cluster_3_voting_5000_rep_1655$Dated <- dated
df_cluster_5_voting_5000_rep_1655$Dated <- dated
#
df_cluster_voting_all <- rbind(
  df_cluster_voting_all,
  df_cluster_3_voting_5000_rep_1655,
  df_cluster_5_voting_5000_rep_1655
)
#
df_cluster_3_row_voting_5000_rep_1655$Dated <- dated
df_cluster_5_row_voting_5000_rep_1655$Dated <- dated
#
df_cluster_row_voting_all <- rbind(
  df_cluster_row_voting_all,
  df_cluster_3_row_voting_5000_rep_1655,
  df_cluster_5_row_voting_5000_rep_1655
)
```

Archive  

```{r }
save(
  df_cluster_3_voting_5000_rep_1655,
  df_cluster_5_voting_5000_rep_1655,
  df_cluster_3_row_voting_5000_rep_1655,
  df_cluster_5_row_voting_5000_rep_1655,
  file=paste0(output_dir,"archive_cluster_voting_5000_rep_1655_all.RData")
)
```  

KEY STEP    

```{r }
str(df_cluster_voting_all)
str(df_cluster_row_voting_all)
#
save(
  df_cluster_voting_all,
  df_cluster_row_voting_all,
  file="cluster_voting_all.RData"
)
```

Archive  

```{r }
save(
  df_cluster_5_voting_5000_rep_1655,
  df_cluster_5_row_voting_5000_rep_1655,
  df_cluster_3_voting_5000_rep_1655,
  df_cluster_3_row_voting_5000_rep_1655,
  file=paste0(output_dir,"archive_cluster_voting_5000.RData")
)
```

### ATTEND TO POPULATION SIMS  

All we have is what Jeremy contributed.  

Use df_cluster_pop_all and df_cluster_row_pop_all  

```{r }
dated <- Sys.time()
#
df_cluster_pop_all <- 
  rbind(
    df_cluster_3_jf_10k_rep,
    df_cluster_5_jf_10k_rep
  )
df_cluster_pop_all$Dated <- dated
#
df_cluster_row_pop_all <- 
  rbind(
    df_cluster_3_row_jf_10k_rep,
    df_cluster_5_row_jf_10k_rep
  )
df_cluster_row_pop_all$Dated <- dated
```

Archive  

```{r }
save(
  df_cluster_5_jf_10k_rep,
  df_cluster_5_row_jf_10k_rep,
  df_cluster_3_jf_10k_rep,
  df_cluster_3_row_jf_10k_rep,
  file=paste0(output_dir,"archive_cluster_ACS_jf_10k.RData")
)
```  
KEY STEP    

```{r }
str(df_cluster_pop_all)
str(df_cluster_row_pop_all)
#
save(
  df_cluster_pop_all,
  df_cluster_row_pop_all,
  file="cluster_pop_all.RData"
)
```

