---
title: "AreWeRepresented Munging NCVOTER ID Fields"
output:
  html_document: default
author: "D. Hopp"
date: "October 6, 2018"
---

Read the NCSBE ncvoter files (we have only one) and retain only the "id" fields.  

This is a stand-alone, single purpose Rmd script.  

```{r a1}
require(data.table)
```
```{r a2}
data_dir <- "../AWR_data_only/"
data_ncvoter_dir <- paste0(data_dir,"ncvoter/")
data_ncvhis_dir <- paste0(data_dir,"ncvhis/")
data_RData_dir <- paste0(data_dir,"RData/")
```

Retain fields:  

```{r b1} 
v_ncvoter_cols <- c(
  "county_id",
  "county_desc",
  "voter_reg_num",
  "status_cd",
  "reason_cd",
  "last_name",
  "first_name",
  "middle_name",
  "name_suffix_lbl",
  "res_street_address",
  "registr_dt",
  "race_code",
  "ethnic_code",
  "party_cd",
  "gender_code",
  "birth_year",
  "ncid"
)
```
Should be able to do this but it takes a long time:  "unzip -q data/ncvoter_Statewide_180818.zip"  
Has to be serialized. Why bother to do every time this is run?  
Appears that zipped file should have same name as .zip, or that there should be some way to explicitly name the the zipped file we are looking for ... maybe.  
  
Build and save ncvoter.  

```{r c1}
do_build <- TRUE
```  

Do not trust the quality of the data.  
Coerce middle_name of length 1 to add a period. This is for standardization.  

```{r d1}
if (do_build){
  df_ncvoter_subset <- data.table::fread(
    paste0(data_ncvoter_dir,"ncvoter_Statewide_20180407.txt"),
    header=TRUE,
    sep="\t",
    stringsAsFactors=FALSE,
    select=v_ncvoter_cols,
    data.table=FALSE
  ) %>%
  dplyr::mutate(
    last_name=trimws(last_name),
    first_name=trimws(first_name),
    middle_name=trimws(middle_name),
    middle_name=ifelse(
      nchar(middle_name)==1,
      paste0(middle_name,"."),
      middle_name
    ),
    name_suffix_lbl=trimws(name_suffix_lbl),
    middle_init=ifelse(
      nchar(middle_name)>0,
      paste0(stringr::str_sub(middle_name,1,1),"."),
      ""
    ),
    as_of="20180407"
  ) %>%
  dplyr::arrange(
    county_desc,
    last_name,
    first_name,
    middle_name,
    name_suffix_lbl,
    ncid
  )
  #
  str(df_ncvoter_subset)
  #
  save(
    df_ncvoter_subset,
    file=paste0(data_RData_dir,"df_ncvoter_subset_20180818.RData")
  )
}
```  

### Look for duplicate records on basis of all columns except as noted (includes address).  

```{r e1}
N_A <- nrow(
  df_ncvoter_subset %>%
  dplyr::distinct(
    county_id,
    county_desc,
    status_cd,
    reason_cd,
    last_name,
    first_name,
    middle_name,
    middle_init,
    name_suffix_lbl,
    res_street_address,
    registr_dt,
    race_code,
    ethnic_code,
    party_cd,
    gender_code,
    birth_year,
    as_of
  )
)
```

Shrinkage by `r nrow(df_ncvoter_subset)-N_A` rows, to about `r round(N_A*100/nrow(df_ncvoter_subset),digits=1)`%.  

Note: ncid distinguishes otherwise identical records.  

### Look for duplicate records on basis of all columns except as noted (does NOT include address).  

```{r f1}
df_ncvoter_subset_distinct_demog <-
  df_ncvoter_subset %>%
  dplyr::distinct(
    county_id,
    county_desc,
    last_name,
    first_name,
    middle_name,
    middle_init,
    name_suffix_lbl,
    race_code,
    ethnic_code,
    party_cd,
    gender_code,
    birth_year,
    as_of
  )
#
str(df_ncvoter_subset_distinct_demog) 
```  

Shrinkage by `r nrow(df_ncvoter_subset)-nrow(df_ncvoter_subset_distinct_demog)` rows, to about `r round(nrow(df_ncvoter_subset_distinct_demog)*100/nrow(df_ncvoter_subset),digits=1)`%.  

```{r g1}
save(
  df_ncvoter_subset,
  file=paste0(data_RData_dir,"df_ncvoter_subset_20180407.RData")
)
save(
  df_ncvoter_subset_distinct_demog,
  file=paste0(data_RData_dir,"df_ncvoter_subset_distinct_demog_20180407.RData")
)
```  
